### 2021-11-04 Tables/studentOutVarDescr.tex
methods results re-render

\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:studentOutVarDescr}Student: Multilevel Core Variable Descriptives (Outgroup Contact Only)}
\centering
\begin{tabular}[t]{llccccc}
\toprule
  & Core Need & Competence & Autonomy & Relatedness & Quality & Attitudes NL\\
\midrule
Core Need &  & 0.44** & 0.49*** & -0.22 & 0.51*** & 0.14\\
Competence & 0.23*** &  & 0.58*** & 0.40** & 0.55*** & 0.21\\
Autonomy & 0.12*** & 0.24*** &  & 0.22 & 0.61*** & 0.05\\
Relatedness & 0.48*** & 0.31*** & 0.38*** &  & 0.35* & -0.09\\
Quality & 0.19*** & 0.40*** & 0.16*** & 0.15*** &  & 0.16\\
Attitudes NL & 0.32*** & 0.36*** & 0.19*** & 0.23*** & 0.30*** & \\
\addlinespace
Grand Mean & 86.86 & 73.23 & 78.58 & 60.30 & 78.80 & 70.41\\
Between-person SD & 11.20 & 13.95 & 14.07 & 17.35 & 10.71 & 17.13\\
Within-person SD & 15.87 & 16.81 & 14.24 & 26.14 & 17.88 & 9.87\\
ICC(1) & 0.14 & 0.27 & 0.40 & 0.19 & 0.14 & 0.72\\
ICC(2) & 0.58 & 0.76 & 0.85 & 0.67 & 0.59 & 0.96\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Upper triangle: Between-person correlations;}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Lower triangle: Within-person coorelations;}\\
\multicolumn{7}{l}{\rule{0pt}{1em}*** p < .001, ** p < .01,  * p < .05}\\
\end{tabular}
\end{minipage}
\end{table}