### 2021-12-16 Tables/mdlTheory.tex

\begin{landscape}\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:intergroupNeedsTbl}Core Need Fulfillment and Allport's Conditions}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{ccccccccccccccccccccccccccc}
\toprule
\multicolumn{1}{c}{} & \multicolumn{6}{c}{Study 1 - CoreNeed} & \multicolumn{6}{c}{Study 2 - CoreNeed} & \multicolumn{6}{c}{Study 3 - CoreNeed} & \multicolumn{6}{c}{Study 3 - Allport} & \multicolumn{2}{c}{Study 3 - CoreNeed + Allport} \\
\cmidrule(l{3pt}r{3pt}){2-7} \cmidrule(l{3pt}r{3pt}){8-13} \cmidrule(l{3pt}r{3pt}){14-19} \cmidrule(l{3pt}r{3pt}){20-25} \cmidrule(l{3pt}r{3pt}){26-27}
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Quality } & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Quality } & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Quality } & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Quality } & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Attitude} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-15} \cmidrule(l{3pt}r{3pt}){16-17} \cmidrule(l{3pt}r{3pt}){18-19} \cmidrule(l{3pt}r{3pt}){20-21} \cmidrule(l{3pt}r{3pt}){22-23} \cmidrule(l{3pt}r{3pt}){24-25} \cmidrule(l{3pt}r{3pt}){26-27}
 & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$\\
\midrule
\addlinespace[0.3em]
\multicolumn{27}{l}{\textbf{Fixed}}\\
\hspace{1em}(Intercept) & 24.28**** [20.20, 28.36] &  & 72.53**** [66.83, 78.24] &  & 72.50**** [66.75, 78.24] &  & 78.84**** [76.86, 80.83] &  & 70.71**** [67.55, 73.87] &  & 70.67**** [67.50, 73.84] &  & 81.36**** [78.53, 84.19] &  & 68.32**** [65.10, 71.54] &  & 68.32**** [65.09, 71.54] &  & 81.42**** [78.57, 84.26] &  & 68.36**** [65.16, 71.56] &  & 68.34**** [65.12, 71.56] &  & 68.33**** [65.05, 71.61] & \\
\hspace{1em}CoreNeed & 0.42**** [ 0.32,  0.52] & 0.40 [ 0.31, 0.49] & 0.17** [ 0.06,  0.29] & 0.30 [ 0.18, 0.41] & 0.04 [-0.08,  0.15] & 0.13 [ 0.00, 0.26] & 0.40**** [ 0.28,  0.52] & 0.27 [ 0.18, 0.37] & 0.13*** [ 0.07,  0.19] & 0.14 [ 0.06, 0.22] & 0.03 [-0.01,  0.07] & 0.07 [-0.01, 0.15] & 0.43**** [ 0.33,  0.54] & 0.31 [ 0.24, 0.38] & 0.19**** [ 0.12,  0.27] & 0.18 [ 0.11, 0.24] & 0.11** [ 0.05,  0.17] & 0.11 [ 0.05,  0.17] &  &  &  &  &  &  & 0.13**** [ 0.08,  0.17] & 0.12 [ 0.05, 0.18]\\
\hspace{1em}Quality &  &  &  &  & 0.23*** [ 0.12,  0.33] & 0.41 [ 0.29, 0.54] &  &  &  &  & 0.17**** [ 0.11,  0.23] & 0.24 [ 0.14, 0.34] &  &  &  &  & 0.17**** [ 0.12,  0.22] & 0.17 [ 0.12,  0.22] &  &  &  &  & 0.17**** [ 0.12,  0.23] & 0.22 [ 0.15, 0.29] &  & \\
\hspace{1em}Allport &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  & 0.65**** [ 0.54,  0.76] & 0.46 [ 0.39, 0.52] & 0.22**** [ 0.15,  0.29] & 0.21 [ 0.14, 0.27] & 0.09** [ 0.03,  0.15] & 0.10 [ 0.03, 0.16] & 0.17*** [ 0.09,  0.24] & 0.17 [ 0.11, 0.23]\\
\addlinespace[0.3em]
\multicolumn{27}{l}{\textbf{Random}}\\
\hspace{1em}$\sigma^2$ & 261.1 &  & 63.1 &  & 37.61 &  & 275.7 &  & 89.83 &  & 76.47 &  & 222.5 &  & 91.84 &  & 82.15 &  & 188.9 &  & 93.05 &  & 83.82 &  & 90.78 & \\
\hspace{1em}$\tau_{00}$ & 71.13 (PID) &  & 172.54 (PID) &  & 177.20 (PID) &  & 63.00 (PID) &  & 259.79 (PID) &  & 264.19 (PID) &  & 131.33 (PID) &  & 182.38 (PID) &  & 183.10 (PID) &  & 134.62 (PID) &  & 179.74 (PID) &  & 182.79 (PID) &  & 189.33 (PID) & \\
\hspace{1em}$\tau_{11}$ &  &  & 0.05 (PID.CoreNeed) &  & 0.05 (PID.CoreNeed)\n0.04 (PID.Quality) &  & 0.12 (PID.CoreNeed) &  & 0.02 (PID.CoreNeed) &  & 0.00 (PID.CoreNeed)\n0.04 (PID.Quality) &  & 0.09 (PID.CoreNeed) &  & 0.05 (PID.CoreNeed) &  & 0.03 (PID.CoreNeed)\n0.02 (PID.Quality) &  & 0.11 (PID.Allport) &  & 0.05 (PID.Allport) &  & 0.02 (PID.Allport)\n0.02 (PID.Quality) &  & 0.00 (PID.CoreNeed)\n0.04 (PID.Allport) & \\
\hspace{1em}$\rho_{01}$ &  &  & 0.03 (objVar$cor.slope_intercept.) &  & 0.37 (PID.CoreNeed)\n-0.22 (PID.Quality) &  & -0.09 (objVar$cor.slope_intercept.) &  & -0.16 (objVar$cor.slope_intercept.) &  & -0.18 (PID.CoreNeed)\n-0.06 (PID.Quality) &  & -0.26 (objVar$cor.slope_intercept.) &  & -0.26 (objVar$cor.slope_intercept.) &  & -0.14 (PID.CoreNeed)\n-0.41 (PID.Quality) &  & -0.07 (objVar$cor.slope_intercept.) &  & -0.19 (objVar$cor.slope_intercept.) &  & 0.12 (PID.Allport)\n-0.51 (PID.Quality) &  & -1.00 (PID.CoreNeed)\n-0.07 (PID.Allport) & \\
\hspace{1em}$\rho_{00}$ &  &  &  &  & -0.47 (PID.CoreNeed-Quality) &  &  &  &  &  & 0.79 (PID.CoreNeed-Quality) &  &  &  &  &  & 0.10 (PID.CoreNeed-Quality) &  &  &  &  &  & 0.11 (PID.Allport-Quality) &  & 0.07 (PID.CoreNeed-Allport) & \\
\addlinespace[0.3em]
\multicolumn{27}{l}{\textbf{Fit}}\\
\hspace{1em}AIC & 3294 &  & 2816 &  & 2672 &  & 8073 &  & 7222 &  & 7141 &  & 14054 &  & 12632 &  & 12494 &  & 13805 &  & 12652 &  & 12519 &  & 12625 & \\
\hspace{1em}BIC & 3309 &  & 2839 &  & 2712 &  & 8102 &  & 7251 &  & 7190 &  & 14087 &  & 12665 &  & 12548 &  & 13838 &  & 12684 &  & 12573 &  & 12679 & \\
\hspace{1em}N & 387 &  & 387 &  & 387 &  & 935 &  & 935 &  & 935 &  & 1676 &  & 1672 &  & 1672 &  & 1676 &  & 1672 &  & 1672 &  & 1672 & \\
\hspace{1em}Groups & 21 &  & 21 &  & 21 &  & 108 &  & 108 &  & 108 &  & 70 &  & 70 &  & 70 &  & 70 &  & 70 &  & 70 &  & 70 & \\
\hspace{1em}ICC & 0.2141 &  & 0.7447 &  & 0.8399 &  & 0.2547 &  & 0.7471 &  & 0.7835 &  & 0.3933 &  & 0.6738 &  & 0.7006 &  & 0.4433 &  & 0.6679 &  & 0.6963 &  &  & \\
\hspace{1em}$R^2_{marginal}$ / $R^2_{conditional}$ & 0.117 / 0.306 &  & 0.030 / 0.752 &  & 0.067 / 0.851 &  & 0.102 / 0.331 &  & 0.012 / 0.750 &  & 0.028 / 0.789 &  & 0.070 / 0.436 &  & 0.019 / 0.680 &  & 0.041 / 0.713 &  & 0.149 / 0.526 &  & 0.024 / 0.676 &  & 0.040 / 0.709 &  & 0.086 / NA & \\
\bottomrule
\multicolumn{27}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{27}{l}{\rule{0pt}{1em}PID = Participant ID number (grouping variable)}\\
\multicolumn{27}{l}{\rule{0pt}{1em}\$\textbackslash{}sigma\textasciicircum{}2\$: residual variance, \$\textbackslash{}tau\_\{00\}\$: random intercept variance, \$\textbackslash{}tau\_\{11\}\$: random slope variance(s), \$\textbackslash{}rho\_\{01\}\$: random slope-intercept correlation(s) , \$\textbackslash{}rho\_\{00\}\$: correlation(s) between random slopes}\\
\multicolumn{27}{l}{\rule{0pt}{1em}**** \$p\$ < .0001, *** \$p\$ < .001, ** \$p\$ < .01, * \$p\$ < .05}\\
\end{tabular}}
\end{minipage}
\end{table}
\end{landscape}