### 2021-12-16 Tables/mdlContactGeneral.tex
re-run

\begin{landscape}\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:intergroupGeneralTbl}Intergroup General}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccccccccccc}
\toprule
\multicolumn{1}{c}{} & \multicolumn{6}{c}{Aggregated [across participants]} & \multicolumn{6}{c}{Contemporaneous [within participants]} \\
\cmidrule(l{3pt}r{3pt}){2-7} \cmidrule(l{3pt}r{3pt}){8-13}
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Study 1} & \multicolumn{2}{c}{Study 2} & \multicolumn{2}{c}{Study 3} & \multicolumn{2}{c}{Study 1} & \multicolumn{2}{c}{Study 2} & \multicolumn{2}{c}{Study 3} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13}
 & $B$ & $\eta_p$ & $B$ & $\eta_p$ & $B$ & $\eta_p$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$\\
\midrule
(Intercept) & 70.50**** [66.06, 74.94] &  & 67.62**** [64.16, 71.08] &  & 64.83**** [61.50, 68.16] &  & 67.44**** [61.71, 73.17] &  & 63.72**** [59.42, 68.01] &  & 56.30**** [51.58, 61.02] & \\
SumContactNL & 0.27 [-0.15,  0.69] & 0.02 & 0.81** [ 0.32,  1.29] & 0.08 & 0.15 [-0.08,  0.38] & 0.02 &  &  &  &  &  & \\
AvQuality & 0.84** [ 0.36,  1.33] & 0.39 & 0.25 [-0.14,  0.64] & 0.01 & 0.36* [ 0.04,  0.68] & 0.07 &  &  &  &  &  & \\
SumContactNL:AvQuality & -0.05** [-0.08, -0.01] & 0.34 & 0.02 [-0.03,  0.08] & 0.01 & 0.01 [-0.02,  0.03] & 0.01 &  &  &  &  &  & \\
OutgroupInteraction &  &  &  &  &  &  & 2.48**** [ 1.36,  3.59] & 1.07 [ 0.59,  1.54] & 2.99*** [ 1.45,  4.54] & 0.11 [ 0.07, 0.16] & 5.63**** [ 3.97,  7.29] & 0.21 [ 0.15, 0.27]\\
NonOutgroupInteraction &  &  &  &  &  &  & 0.43 [-0.68,  1.54] & 0.16 [-0.32,  0.63] & 0.01 [-0.76,  0.78] & 0.01 [-0.03, 0.05] & 0.34 [-0.49,  1.18] & 0.02 [-0.02, 0.05]\\
\addlinespace[0.3em]
\multicolumn{13}{l}{\textbf{Random}}\\
\hspace{1em}$\sigma^2$ &  &  &  &  &  &  & 69.92 &  & 77.46 &  & 88.1 & \\
\hspace{1em}$\tau_{00}$ &  &  &  &  &  &  & 164.13 (PID) &  & 488.38 (PID) &  & 366.51 (PID) & \\
\hspace{1em}$\tau_{11}$ &  &  &  &  &  &  &  &  & 48.94 (PID.OutgroupInteraction)\n 7.23 (PID.NonOutgroupInteraction) &  & 40.53 (PID.OutgroupInteraction)\n 3.43 (PID.NonOutgroupInteraction) & \\
\hspace{1em}$\rho_{01}$ &  &  &  &  &  &  &  &  & -0.72 (PID.OutgroupInteraction)\n 0.12 (PID.NonOutgroupInteraction) &  & -0.70 (PID.OutgroupInteraction)\n-0.28 (PID.NonOutgroupInteraction) & \\
\hspace{1em}$\rho_{00}$ &  &  &  &  &  &  &  &  & 0.00 (PID.OutgroupInteraction-NonOutgroupInteraction) &  & 0.26 (PID.OutgroupInteraction-NonOutgroupInteraction) & \\
\addlinespace[0.3em]
\multicolumn{13}{l}{\textbf{Fit}}\\
\hspace{1em}Groups &  &  &  &  &  &  & 23 &  & 113 &  & 71 & \\
\hspace{1em}N & 21 &  & 113 &  & 71 &  & 1225 &  & 4965 &  & 3907 & \\
\hspace{1em}F Test & F(3, 17) = 6.66, p =  0.004 &  & F(3, 109) = 3.75, p =  0.013 &  & F(3, 67) = 2.15, p =  0.102 &  &  &  &  &  &  & \\
\hspace{1em}$R^{2 (a)}$ & 0.540 / 0.459 &  & 0.094 / 0.069 &  & 0.088 / 0.047 &  & 0.006 / 0.703 &  & 0.003 / 0.817 &  & 0.026 / 0.708 & \\
\hspace{1em}AIC &  &  &  &  &  &  & 8788 &  & 36476 &  & 29074 & \\
\hspace{1em}BIC &  &  &  &  &  &  & 8814 &  & 36542 &  & 29137 & \\
\hspace{1em}ICC &  &  &  &  &  &  & 0.7013 &  & 0.8167 &  & 0.7007 & \\
\bottomrule
\multicolumn{13}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{13}{l}{\rule{0pt}{1em}PID = Participant ID number (grouping variable)}\\
\multicolumn{13}{l}{\rule{0pt}{1em}$sigma^2$: residual variance, $tau_{00}$: random intercept variance, $tau_{11}$: random slope variance(s), $rho_{01}$: random slope-intercept correlation(s) , $rho_{00}$: correlation(s) between random slopes}\\
\multicolumn{13}{l}{\rule{0pt}{1em}**** $p$ < .0001, *** $p$ < .001, ** $p$ < .01, * $p$ < .05}\\
\multicolumn{13}{l}{\rule{0pt}{1em}\textsuperscript{a} For linear regression: $R^2 / R^2_{adjusted}$, for multilevel regression: $R^2_{marginal} / R^2_{conditional}$ }\\
\end{tabular}}
\end{minipage}
\end{table}
\end{landscape}