### 2022-01-04 Tables/workerOutVarDescr.tex

\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:workerOutVarDescr}Worker: Multilevel Core Variable Descriptives (Outgroup Contact Only)}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llccccc}
\toprule
  & Core Need & Competence & Autonomy & Relatedness & Quality & Attitudes NL\\
\midrule
Core Need &  & 0.82*** & 0.60*** & 0.32 & 0.52** & -0.03\\
Competence & 0.36*** &  & 0.89*** & 0.24 & 0.38 & -0.24\\
Autonomy & 0.28*** & 0.23*** &  & 0.29 & 0.56** & 0.02\\
Relatedness & 0.51*** & 0.39*** & 0.37*** &  & -0.09 & 0.14\\
Quality & 0.17*** & 0.44*** & 0.27*** & 0.27*** &  & 0.50*\\
Attitudes NL & 0.24*** & 0.37*** & 0.24*** & 0.37*** & 0.52*** & \\
\addlinespace
Grand Mean & 32.20 & 16.53 & 22.64 & 1.58 & 24.10 & 72.46\\
Between SD & 12.42 & 13.41 & 10.22 & 14.16 & 9.50 & 13.62\\
Within SD & 17.66 & 18.78 & 14.93 & 23.97 & 18.85 & 9.50\\
ICC(1) & 0.33 & 0.31 & 0.30 & 0.21 & 0.18 & 0.68\\
ICC(2) & 0.90 & 0.89 & 0.89 & 0.83 & 0.80 & 0.98\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Upper triangle: Between-person correlations;}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Lower triangle: Within-person correlations;}\\
\multicolumn{7}{l}{\rule{0pt}{1em}*** p < .001, ** p < .01,  * p < .05}\\
\end{tabular}}
\end{minipage}
\end{table}