### 2022-11-04 manuscript.tex
general intro comments adjusted

% define document type (i.e., template. Here: A4 APA manuscript with 12pt font)
\documentclass[man, 12pt, a4paper, mask]{apa7}

% change margins (e.g., for margin comments):
%\usepackage{geometry}
% \geometry{
% a4paper,
% marginparwidth=30mm,
% right=50mm,
%}

% add packages
\usepackage[american]{babel}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{hyperref}
\usepackage[style=apa, sortcites=true, sorting=nyt, backend=biber, natbib=true, uniquename=false, uniquelist=false, useprefix=true]{biblatex}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{setspace,caption}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{lipsum}
\usepackage{soul}
\usepackage{xcolor}
\usepackage{fourier}
\usepackage{stackengine}
\usepackage{scalerel}
\usepackage{fontawesome5}
\usepackage[normalem]{ulem}
% \usepackage{longtable}
\usepackage{amsmath}
\usepackage{ntheorem}
\usepackage{afterpage}
\usepackage{float}
\usepackage{array}
\usepackage{censor}
\usepackage{pdflscape}
\usepackage{lscape}
\usepackage{pdfpages}
\usepackage{enumitem}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{makecell}
\usepackage{tabu}
% Path Diagrams
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}
\tikzset{mynode/.style={draw,text width=1in,align=center} }
\usetikzlibrary{positioning}

% make warning with red triangle
\newcommand\Warning[1][2ex]{%
  \renewcommand\stacktype{L}%
  \scaleto{\stackon[1.3pt]{\color{red}$\triangle$}{\tiny\bfseries !}}{#1}}%

% make question with red triangle
\newcommand\Question[1][2ex]{%
  \renewcommand\stacktype{L}%
  \scaleto{\stackon[1.3pt]{\color{red}$\triangle$}{\tiny\bfseries ?}}{#1}}%

% add definition sections
\theoremstyle{break}
\newtheorem{definition}{Definition}

% add hypothesis sections
\theoremstyle{plain}
\theoremseparator{:}
\newtheorem{hyp}{Hypothesis}

\newtheorem{subhyp}{Hypothesis}
   \renewcommand\thesubhyp{\thehyp\alph{subhyp}}

% Analysis Plan sections (probably better as Definitions or Remarks instead of Theorems)
\newtheorem{ap}{Hypothesis}

\newtheorem{subap}{Hypothesis}
   \renewcommand\thesubap{\theap\alph{subap}}

% add quote section
\usepackage{csquotes}

% framed box section
\usepackage{framed}
\emergencystretch=1em

% formatting links in the PDF file
\hypersetup{
pdfpagemode={UseOutlines},
bookmarksopen=true,
bookmarksopenlevel=0,
hypertexnames=false,
colorlinks   = true, %Colours links instead of ugly boxes
urlcolor     = blue, %Colour for external hyperlinks
linkcolor    = blue, %Colour of internal links
citecolor   = cyan, %Colour of citations
pdfstartview={FitV},
unicode,
breaklinks=true,
}

% custom open science badge height
\newlength{\badgeheight}
\setlength{\badgeheight}{1em}

% language settings
\DeclareLanguageMapping{american}{american-apa}

% add reference library file
\addbibresource{references.bib}

% Title and header
\title{Psychological Needs During Intergroup Contact: Three Experience Sampling Studies}
\shorttitle{Needs in Intergroup Contact}

% Authors
\author[*,1]{Jannis Kreienkamp}
\author[1]{Maximilian Agostini}
\author[1]{Laura F. Bringmann}
\author[1]{Peter de Jonge}
\author[1]{Kai Epstude}
\affiliation{\hfill}

\affil[1]{University of Groningen, Department of Psychology}

\authornote{
   \addORCIDlink{* Jannis Kreienkamp}{0000-0002-1831-5604}\\
   \addORCIDlink{Maximilian Agostini}{0000-0001-6435-7621}\\
   \addORCIDlink{Laura F. Bringmann}{0000-0002-8091-9935}\\
   \addORCIDlink{Peter de Jonge}{0000-0002-0866-6929}\\
   \addORCIDlink{Kai Epstude}{0000-0001-9817-3847}

We have no known conflict of interest to declare. The authors received no specific funding for this work. Materials and software are available at \url{https://janniscodes.github.io/intergroup-contact-needs/}  \citep{Kreienkamp2022}. Protocols, materials, data, and code are available at \url{https://osf.io/pr9zs/?view_only=208a53a1f0ff48dda1c17357328fa578} \citep{Kreienkamp2022a}. The preregistration of Study 3 can be accessed as part of our Open Science Framework repository \citep{Kreienkamp2021f}

Correspondence concerning this article should be addressed to Jannis Kreienkamp, Department of Psychology, University of Groningen, Grote Kruisstraat 2/1, 9712 TS Groningen (The Netherlands).  E-mail: j.kreienkamp@rug.nl}

\leftheader{Kreienkamp}

% Abstract
\abstract{
\hl{needs to be adjusted}\\
One challenge of modern intergroup contact research has been the question of when and why an interaction is perceived as positive and leads to better intergroup relations. We propose to consider situational psychological needs during everyday intergroup contact. We conducted three extensive longitudinal studies with recent migrants, to capture their interactions with the majority outgroup (total N of measurements = 10,297). \hl{We find inconsistent results for a traditional test of the contact hypothesis with experience sampling data. However, when using multilevel models to test the contact hypothesis, we find that individual outgroup interactions and perceived interaction quality have a strong influence on positive outgroup attitudes.} 
The fulfillment of psychological needs is consistently a strong predictor of positive attitudes during intergroup interactions and these positive effects emerged via perceived interaction quality. The situational needs remained the core predictor of outgroup attitudes even after controlling for Allport’s (\citeyear{Allport1954b}) contact conditions or more general self-determination needs \citep[][]{Ryan2017}. As one of the first studies to test intergroup contact theory using extensive longitudinal data, we offer insight into the mechanisms of positive intergroup contact during real-life interactions and find situational motivations to be a key building block of understanding and addressing positive intergroup interactions.

\noindent\textbf{Public significance statement}: In this paper, we provide evidence that the fulfillment of situational psychological needs during real-life intergroup contacts meaningfully predicts perceived interaction quality and positive outgroup attitudes. Methodologically, this offers testament to the emerging practice of capturing real-life interactions using extensive longitudinal data. Theoretically, our results give weight to motivational fulfillment as a flexible and effective mechanism in understanding positive intergroup contact.
}

\keywords{
    Intergroup Contact, Psychological Needs, Outgroup Attitudes, Interaction Quality, Extensive Longitudinal Data\\
    \textit{Open Science Practices:}
    \noindent \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/registrationplus-color.png}} Preregistration+, 
    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/material-color.png}} Open Materials, 
    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/data-color.png}} Open Data,\break
    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/code-color.png}} Open Code, 
    \href{https://osf.io}{\includegraphics[height=\badgeheight]{assets/open-badges-small/supplements-color.png}} Open Supplements
    }

% set indentation size
\setlength\parindent{1.27cm}

% Start of the main document:
\begin{document}

% add title information (incl. title page and abstract)
\maketitle

% **CHEAT SHEET / LEGEND**
%
% Comments:
% '%' starts a comment in LaTeX (not printed)
% '\todo[inline]{} makes orange boxes in PDF
% '\marginpar{}' notes in margins
% '\footnote{}' footnote
% '\Warning' important note indicator in PDF (triangle with exclamation mark)
% '\Question' question note indicator in PDF (triangle with question mark)
%
% Citation (with Natbib citation style):
% '\citep[e.g.][p. 15]{CitationKey}' citation in parentheses "(e.g., Berry, 2003, p. 15)"
% '\citet{CitationKey}' citation in text "Berry (2003)"
% '\citealt' and '\citealp' alternate citation without parentheses
% '\citeauthor' and '\citeyear' only year or author
% 
% Headings:
% '\part{}' and '\chapter{}' only relevant for multi-part or multi-chapter documents
% '\section{}' heading level 1
% '\subsection{}' heading level 2
% '\subsubsection{}' heading level 3
% '\paragraph{}' heading level 4
% '\subparagraph{}' heading level 5
%
% formatting:
% '\textbf{}' text bold font
% '\textit{}' text italic font
% '\underline{}' text underline
% '\sout{}' text strike out
% '\textsc{}' text small caps
% '\vspace{1em}' add vertical space
% '\hspace{1em}' add horizontal space
% '\\' new line (i.e., line break)
% '\pagebreak' start new page (i.e., page break)
% '\noindent' do not indent current line (e.g., current paragraph)
% 'begin{center}...end{center}' center text or object
%
% Math mode:
% '$\alpha = .8$' mathematical equation inline
% '$$\hat{y} = b_0 + b_1x$$' mathematical equation in its own line
% '\begin{equation}...\end{equation}' multi-line equation
% '\approx' approximate symbol
% '\neq' not equal
% '\bar' mean bar over letter
% '\pm' plus minus sign 
% '^{}' superscript
% '_{}' subscript
% '\fraq{numerator}{denominator}' fraction
% '\sqrt[n]{}' square root
% '\sum_{k=1}^n' sum for 1 through n
%
% Insert things from elsewhere:
% '\input{filename}' inputs the raw (tex) file as a command (e.g., tables and R-Markdown imports)
% '\include{filename}' includes section on new page (incl. possible auxiliary info)
% '\includegraphics[settings]{filename}' add a figure or graph
% '\caption{}' adds a caption to a table or figure
% '\label{}' labels sections, tables, figures, etc. so that they can be referred to.
% '\ref{}' refer to a labelled sections, tables, figures, etc.
% '\begin{enumerate}...\end{enumerate}' numbered list
% '\begin{itemize}...\end{itemize}' bullet-ed list
% '\item' item in list section 
%
% Symbols:
% '\&' and sign
% '\%' percent sign
% '\_' three dotes
% '\#' hash symbol
% ------------------------------------------------------------------

% Migrant Example Pathways
% Relevance (Version 1 of 2): Migrant Example Version:
One of the main intergroup societal issues to date, are the struggles of many migrants across the world, hoping to build a positive relationship with the majority group. The intergroup contact hypothesis postulates that prejudice can be reduced and favorable attitudes be increased if members of two groups have frequent and positive contact \citep[e.g.,][]{Allport1954b, Hewstone1996, Pettigrew1998}. Over the past 70 years, a plethora of studies and interventions have shown the general effectiveness of positive intergroup contact \citep[e.g.,][]{Pettigrew2006}. However, even though a central assumption of intergroup contact theory is that the contact should be positive, relatively little research has thus far explained when and why people perceive their everyday intergroup interactions as positive.

% Full Migrant Pathway
% Relevance (Version 2 of 2): Full Migrant Focus Version:
%The adaptation of migrants in new cultural contexts has become an important issue for many societies around the world. A major aspect of such migrant adaptation arguably unfolds during the daily interactions migrants have with the cultural majority members \citep{Maxwell2017, Sam2010}. One of the main social psychological theories aimed at understanding contact between social groups is ’intergroup contact hypothesis’. In its most essential interpretation, the intergroup contact hypothesis postulates that prejudice can be reduced and favorable attitudes increased if members of two groups have frequent and positive contact \citep[e.g.,][]{Allport1954b, Hewstone1996, Pettigrew1998}. Over the past 70 years, a plethora of studies and interventions have shown the general effectiveness of positive intergroup contact \citep[e.g.,][]{Pettigrew2006}. However, even though a central assumption of intergroup contact theory has been that the contact should be positive, relatively little research has thus far explained when and why people perceive their everyday inter-group interactions as positive.

% Old Relevance: Conflict is a problem, positive contact as solution, but little research on when daily contact is perceived positive
%Conflict between social groups and their individual members remains a prevalent feature of the modern human condition. Experiences of prejudices, discrimination, and animosities with other groups continue to plague the everyday lives of many people around the world. One of the main ameliorations proposed by social psychologists has been the intergroup contact hypothesis. In its most essential interpretation, the intergroup contact hypothesis postulates that frequent and positive contact with an out-group reduces prejudice and increases favorable attitudes towards the other group \citep[e.g.,][]{Allport1954b, Hewstone1996, Pettigrew1998}. And even though over the past 70 years, a plethora of studies and interventions have shown the general effectiveness of positive intergroup contact \citep[e.g.,][]{Pettigrew2006}, relatively little research has thus far explained when and why people's everyday inter-group interactions are perceived as positive.

% Problem v.02: theoretical interaction quality central to understanding outcomes, practical many impactful negative interactions in everyday life
% Problem Illustration Paragraph:       
Importantly, as we still fail to understand when and why an interaction is perceived as positive, substantial theoretical and practical challenges remain. There is now consistent evidence that negative intergroup contacts lead to worse attitudes, prejudice, and reduced future interaction motivation \citep[e.g.,][]{Barlow2012, Prati2021, Graf2014}. In light of these findings, understanding interaction quality thus sits at the heart of understanding when an intergroup contact is successful (e.g., Brown et al. \citeyear{Brown2007}, Tropp et al. \citeyear{Tropp2016}; also cf. Allport’s original sentiment, Allport \citeyear{Allport1954b}). Practically, policymakers and practitioners are thus far often under-prepared to deal with the occurrences of negative interactions, especially in everyday life contexts. Understanding the psychological mechanisms of when and why interactions are perceived as positive is, thus, an important issue for understanding whether an interaction leads to better intergroup perceptions, especially during everyday interactions outside the lab.

% Aim / Solution v.02: look at need fulfillment as a mechanism in daily interactions of migrants
% Proposal Paragraph:
We propose that the key to understanding how an interaction is perceived is to examine the level of need fulfillment it provided to an individual. As an example, if someone seeks acceptance by their interaction partner, and this need is fulfilled during the interaction, the person should rate the interaction and the group of the interaction partner more favorably. To test this, we collected three sets of real-life data from recent immigrants, assessing their daily interactions with majority group members, tracking situational needs, interaction quality, and outgroup attitudes.

\section{Need Mechanism in Intergroup Contact}
% Motivational mechanism:
% interaction quality is important
% past research has either focused on conditions or cognitive-affective processes
% neither are necessary nor explain why an interaction is positive
Several meta-analytic reviews have found that positive intergroup contact reduces prejudice and increases positive attitudes in experimental and cross-sectional studies \citep[][]{Tropp2005, Pettigrew2006, Davies2011}, as well as in intergroup contact interventions outside the lab \citep[][]{Beelmann2014, Lemmer2015}. It is widely accepted that equal status, common goals, collaboration, and structural support during the interaction form the optimal conditions for such positive contact effects \citep[][]{Allport1954b, Pettigrew1969}. However, despite their prominence in guiding research on this topic, meeting the contact conditions does not seem to be essential to finding positive effects of intergroup contact \citep[][]{Pettigrew2006}. It, thus, remains unclear when and why an interaction is perceived as positive and what factors (beyond Allport's conditions) explain positive contact effects\footnote{It should be noted here that following Allport's original conditions, several additional conditions of optimal contact were proposed \citep[for a critical discussion see][]{Pettigrew1986}. Similarly, since the early 21\textsuperscript{st} century many psychological processes during intergroup contact were examined \citep[e.g. see,][]{Paolini2021}. Among others, researchers have explored different forms of social categorizations \citep[][]{Pettigrew1998}, the salience of social categories \citep[][]{Brown2005}, intimacy \citep[e.g.,][]{Marinucci2021} and attachment \citep[e.g.,][]{Tropp2021}, threat and intergroup anxiety \citep[e.g.,][]{Stephan2008, Paolini2004}, and to a lesser extent knowledge about the other group \citep[][]{Pettigrew2008c}. However, all these advances share the underlying criticism that they are not geared towards understanding when and why an interaction is perceived as positive.}. We propose that the literature on motivation and psychological needs may help us understand when and why exactly an interaction is perceived as positive.

% ALTERNATIVE: SHORTER BUT NOT FOCUSED ON ALLPORT'S CONDITIONS (WHICH WE TEST LATER)
% Several meta-analytic reviews found that positive intergroup contact reduces prejudice and increases positive attitudes in experimental and cross-sectional studies \citep[][]{Tropp2005, Pettigrew2006, Davies2011}, as well as in intergroup contact interventions outside the lab \citep[][]{Beelmann2014, Lemmer2015}. Past research has either focused on conditions under which interactions tend to have favorable effects \citep[e.g., Allport's optimal conditions;][]{Allport1954b, Pettigrew1969} or has investigated cognitive-affective processes during contact \citep[e.g., group salience, intimacy, attachment, or intergroup anxiety;][]{Brown2005, Marinucci2021, Tropp2021, Paolini2004}. However, neither the contact conditions nor the cognitive-affective processes are particularly well-equipped to explain when and why an interaction will be perceived as positive. We propose that we may look to the literature on motivation and psychological needs to understand when and why exactly an interaction is perceived as positive.

% Important in broader field: needs have been found to be important in broader intergroup relations and interpersonal contact literature.
Although considerations of psychological needs and motivations have remained markedly absent within the intergroup contact literature, the concepts are in no way foreign to the broader topic of intergroup relations. Psychological needs have for example found fruitful utility in reconciliation research \citep[][]{Shnabel2008}, the study of minority's discrimination experiences \citep[][]{Celebi2017}, and in non-intergroup interactions \citep[][]{Downie2008}. 
% On the most general level, research on intergroup conflicts has, for example, shown that addressing the group-specific needs predicts willingness to reconcile with the other group \citep[][]{Shnabel2008}. For the case of refugee migrants in their relationship with a majority group, a recent study found that identity needs will buffer against discrimination experiences and will protect personal health \citep[][]{Celebi2017}. Additionally, within the field of interpersonal relations, we find first evidence that the fulfillment of fundamental psychological needs might predict perceived interaction quality in non-intergroup interactions \citep[][]{Downie2008}. 
% Promising for intergroup contact: Asked for in intergroup contact reviews and first research looked at social change
Even within the field of intergroup contact, the idea of need fulfillment as a psychological mechanism is not entirely novel \citep[e.g.,][]{Dovidio2017}. First empirical research has even found psychological need fulfillment during intergroup contacts to predict support for social change \citep[][]{Hassler2021}. Considering need fulfillment during intergroup contact might therefore offer a promising psychological mechanism to understand perceived interaction quality and positive outgroup attitudes.

% Difficult to test general mechanism: Either consider too many needs (situationally relevant but not feasible) or too few (feasible but not transferable). Solution: ask for main need + rating of main need  
% MAYBE MOVE THIS TO 'The Present Research Section'
One reason why motivational considerations might have remained absent from the intergroup contact literature is that there is an overwhelming number of individual needs or goals that might be relevant to a person during an intergroup interaction. Researchers considering the motivational content would, thus, either test few hyper-specific needs that might not be transferable to other intergroup contexts or they may need to assess a broad range of fundamental psychological needs. To avoid this predicament we propose to start by using adaptive and responsive survey designs that allow a tailored approach based on the participants' inputs \citep[e.g.,][]{Tourangeau2017}. In particular, we propose to ask the participants to report their main goal during the interaction in a short open-ended question (i.e., situation core need) and with reference to their own response the participants can then indicate how much this need was fulfilled during the interaction (i.e., need fulfillment). Such an adaptive approach allows us to take this initial step and test whether situational psychological needs indeed generally predict perceived interaction quality and positive outgroup attitudes independent of need content.

\section{Intergroup Contact in Daily Life}
While we have argued that a need fulfillment mechanism is relevant to intergroup contact generally, its flexible and broad applicability might be ideally suited to address the pressing issue of understanding natural intergroup contacts outside the lab. Investigations of such `real-life interactions' often suffer from the difficulty that past intergroup contact research has either focused on the mechanisms of individual interactions in artificial lab studies or has focused on longer-term recall self-reports of natural interactions \citep[e.g.,][]{Pettigrew2006}. Even with extended intervention studies, the most fine-grained data available is usually limited to pre-post-control designs. Such empirical practices stand, however, in stark contrast to many of the theoretical advances that have focused on the dynamic nature of intergroup relations \citep[e.g.,][]{Pettigrew1998}, as well as the core idea of the contact hypothesis, which was focused on the daily interactions of people \citep[][]{Allport1954b}. As a result, within the intergroup contact field researchers have called for longitudinal \citep[][]{Pettigrew1998, Pettigrew2008, Pettigrew2008b, Pettigrew2011} and real-life experience-sampling data outside the lab \citep[][]{MacInnis2015, McKeown2017, Dixon2005}. Such data would be able to capture real-life interactions that include interaction-specific mechanism information close to the actual experience\footnote{Additionally, such experience-sampling data can be collected close to the intergroup interactions and would, thus, largely mitigate recall biases. Moreover, because data is nested within participants, experience-sampling data often allows capturing large amounts of high-quality data with relatively few participants.}.

% Used to be difficult but technological and methodological developments (e.g., FormR + HLM) → collect large body of intensive longitudinal data
In the past, such data collections were often unfeasible because they were either physically impractical or too expensive. However, recent technological developments allow us to easily collect experience sampling data on mobile devices \citep[e.g.,][]{Keil2020} or using web-based applications \citep[e.g.,][]{Arslan2020}. At the same time, analytical methods for such more complex data have become more readily available, making the analyses more approachable \citep[e.g., see][]{ODonnell2021}. Given these technological and methodological developments, we were able to collect a large amount of real-life data following the daily intergroup interactions of recent migrants with the majority group members.

\section{The Present Research}
The aim of this paper is essentially threefold. First, our study is among the first to test the fundamental tenets of intergroup contact and Allport's conditions in real-life extensive longitudinal data. Second, we test whether the fulfillment of situational psychological needs is meaningfully related to more positive outgroup attitudes following intergroup interactions. And third, we test the role of interaction quality in  the relationship of psychological need fulfillment and outgroup attitudes. Based on this we formulated three main hypotheses:
\begin{enumerate}[label=H\arabic*:]
    \item Based on the most general understanding of the contact hypothesis, an increase in frequency and quality of contact should jointly account for changes in more favorable outgroup attitudes across extensive longitudinal data.
    \item Based on Allport’s optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data.
    \item Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data. We also expect situational need fulfillments to work at least as well as Allport's optimal contact conditions in predicting outgroup attitudes.
\end{enumerate}
Beyond these main hypotheses, we have formulated several sub-hypotheses that form the basis for our analysis plan (these sub-hypotheses also include several robustness checks and are available in Appendix \ref{app:AppendixHypotheses}).

To test our main hypotheses, we collected three independent sets of extensive longitudinal data (Studies 1–3). The full surveys are available in Supplementary Materials A as well as in our open science repository \citep[including a complete codebook, see][]{KreienkampMasked2022a}. Our most comprehensive study (Study 3) was fully preregistered  \citep[available at][]{KreienkampMasked2021f}. All studies received ethical approval from the University Masked for Peer Review. The fully annotated analyses are available in Supplementary Material B.

\section{Analysis Plan}
\begin{center}
    \noindent\rule{8cm}{0.4pt}\\
    \hl{\textit{New Section:}}\\
    \vspace{-0.25cm}
    \noindent\rule{8cm}{0.4pt}
\end{center}

Given the more complex data structure of the extensive longitudinal data, we opted to test most of our hypotheses using multilevel regression models, where the measurement occasions (level 1) were nested within the participants (level 2). Given that more complex dynamic modeling procedures are not necessary to test these more basic hypotheses, we will focus on the contemporaneous effect of the data (e.g., need fulfillment during the intergroup interaction predicts outgroup attitudes following the interaction). Such an approach is also tolerant to missing data and uneven case numbers within participants. As is common with multilevel analyses, we use a hierarchical modeling approach and always report the final model in-text (for the full modeling process see Supplementary Material B).

Statistical power is notoriously difficult to estimate for intensive longitudinal- and multilevel models because of the complex covariance structures. However, while power simulations require researchers to make several key assumptions, a major meta-analysis of the intergroup contact hypothesis estimates the average effect size of intergroup contact to be at around: $r$ = .21 \citep[][]{Pettigrew2006}. For such a relatively small effect size, cross-sectional data collections would need to collect 175 to 212 participants (for correlational- and linear regression analyses with .8 power respectively). Given that even our smallest sample captured 387 intergroup contacts, each of the studies should have abandoned power to detect even smaller effect sizes than the effects commonly found within the literature. Additionally, both the number of participants and the number of measurements are within the highest percentiles of sample sizes found within the broader intensive longitudinal literature \citep[e.g.,][]{AanhetRot2012}.

For our main theory tests we sequentially focus on five sets of models to test and validate our hypotheses:

\begin{enumerate}
    \item \textbf{Contact hypothesis in extensive longitudinal data.} Within the individual studies, we begin by testing the most basic assumption of the intergroup contact hypothesis that outgroup attitudes should be more positive after outgroup interactions but not after non-outgroup interactions. For this we use multi-level regression analyses, predicting outgroup attitudes from outgroup interaction and non-outgroup interaction dummy variables. We also reintroduce the participant means as level two predictors to fully disentangle measurement- (level 1) and participant (level 2) variances.

    Importantly, this initial test of the contact hypothesis in intensive longitudinal data does not consider the influence of interaction quality. This is due to the structural missingness of interaction quality --- where people are only able to report their interaction quality perceptions if they had an interaction. Such structural missingness should not be imputed, which stands in conflict with (multi-level) linear regression analyses that can only be estimated where all variables have valid data (which would only be measurements following interactions). The best way to assess the joint effects of interaction numbers and interaction quality is thus a test on an aggregate level. However, as such an aggregation reduces statistical power to the participant level, we assess the full contact hypothesis in the `embeddedness and stability across studies' section, where we combine the data of the three studies (see below).
    \item \textbf{Core need fulfillment during intergroup contact.} The main proposal of this manuscript has been the assertion that the fulfillment of situation core needs during an interaction will be associated with more positive interaction quality perceptions and, ultimately, more positive outgroup attitudes. Thus, for the main set of analyses, we focus on the reported outgroup interactions only. For each study, we will use multi-level regressions to test the main three assertions of our proposal (mirroring the basic steps of a traditional mediation analysis).
    \begin{enumerate}
        \item Core need fulfillment during outgroup interactions should predict more positive outgroup attitudes. To test this sub-hypothesis, we use a two-level regression analysis of outgroup attitudes, with core need fulfillment as the main predictor variable.
        \item Core need fulfillment during outgroup interactions should also predict higher perceived interaction quality. This sub-hypothesis mirrors the one above so that perceived interaction quality is predicted by core need fulfillment in a multi-level regression analysis.
        \item The effect of core need fulfillment on outgroup attitudes should be reduced when considered together with perceived interaction quality. To test whether perceived interaction quality is indeed the psychological mechanism that explains the variance in outgroup attitudes, we use a multi-level regression analysis of outgroup attitudes with core need fulfillment and perceived interaction quality as the main predictor variables.
    \end{enumerate}
    \item \textbf{Robustness within studies.} To build further confidence in the effect of core need fulfillment during outgroup interactions, we conduct two additional robustness analyses for each study. 
    \begin{enumerate}
        \item To check for the role of alternate psychological needs, we add the fulfillment of self-determination theory needs (i.e., competence, autonomy, and relatedness) to the multi-level regression. We then also compare the model with models that predicts outgroup attitudes from self-determination theory need fulfillments or core need fulfillments only. 
        \item To ensure that the core need fulfillment is outgroup contact specific, we return to the full sample of intensive longitudinal measurements within each study and test whether there is an interaction effect of outgroup contact (vs. no outgroup contact) and core need fulfillment. We expect that the effect of core need fulfillment is specific to outgroup interactions and not merely due to a more need fulfilled life in general.
    \end{enumerate} 
    \item \textbf{Allport's conditions in extensive longitudinal data.} Within the third study we formally measure all of Allport's optimal contact conditions. 
    \begin{enumerate}
        \item We use multi-level regression models to test whether the fulfillment of Allport's conditions in real-life data predicts more positive outgroup attitudes and higher perceived interaction quality, using the same approach as for the core need fulfillment above.
        \item We then compare the effect of Allport's contact conditions with our core need fulfillment by comparing the model fit statistics of the two individual models and by adding both concepts to a joint multi-level regression model, to see whether the two approaches explain the same variance in outgroup attitudes.
    \end{enumerate}
    \item \textbf{Embeddedness and stability across studies.} Within the final set of analyses, we look at the broader picture of our results and leverage the data from all participants to contextualize our results.  
    \begin{enumerate}
        \item We first assess the stability of our main analyses. We use forest plots (including meta analytical estimates) to visualize the direction and effect sizes of our three studies. 
        \item We then combine the data from the three studies to test the basic tenets of intergroup contact hypothesis more broadly within the intensive longitudinal real-life data. As we have mentioned above, assessing the effects of intergroup contacts and perceived interaction quality jointly is difficult within a multi-level regression because quality perceptions are only available when participants had an interaction. To circumvent this issue, we use the data from all participants and aggregate the number of interactions participants had, their average interaction quality perceptions, as well as their average outgroup attitudes. We can then use a linear regression analysis to assess the joint effects of intergroup contacts and interaction quality to predict outgroup attitudes (while controlling for the participants' study). This not only offers a more general test of the contact hypothesis with real-life data, but the aggregated data also mirrors the most common methodology of assessing intergroup contact effects within the cross-sectional literature --- where participants are asked to recall how many interactions they had over a one month period, how positive these interactions were, and what their general attitudes towards the outgroup are. This analysis will thus be able to explore potential discrepancies between our main multi-level analyses, the aggregation of extensive longitudinal data, and traditional participant recall studies.
        \item We, finally, use the qualitative data from the participants' self-identified core needs to contextualize our results. We leverage machine learning to extract a topic model of the free-text entries across the three studies. We describe the extracted topics and themes, and compare them to the need contents usually found and measured within the psychological literature. Full methodological details and visualizations are available in Supplemental Material C.
    \end{enumerate}
\end{enumerate}

% Methods and Results from RMarkdown render
\input{Methods-and-Results}

\section{Discussion}
% aims re-iterated
The aim of this project was threefold. We (1) tested the basic predictions of the intergroup contact hypothesis and Allport's optimal conditions in real-life extensive longitudinal data. As a flexible alternative, we (2) proposed that the fulfillment of situational psychological needs meaningfully predicts positive outgroup attitudes. Finally, we (3) assessed whether the fulfillment of psychological needs have their effect (partially) through perceived interaction quality.

% Contact hypothesis in ESM data [inconsistent in aggregate but consistent in ML analysis]
When considering the results of the three studies jointly, we found mixed results for the basic intergroup contact hypothesis. More specifically, we observe that separating between-participant and within-participant effects is crucial. For between-subjects effects (using aggregated reports), the positive effect of more outgroup contact did not depend on the average interaction quality. These aggregate findings are surprising because most past cross-sectional studies have also considered an aggregate measurement of interaction quality (albeit not one from actual daily reports). The absence of this aggregate effect could underline the fact that cross-sectional retrospective data might be misleading because (1) it presents a mixture of within and between-subjects effects \citep[][]{Hamaker2020}, or (2) suffers from recall biases (e.g., where times with no outgroup interactions are undervalued by participants during the retrospective evaluations).\footnote{It should be noted that such inconsistencies with past research might in part be a data artefact (e.g., because most people reported substantially more measurements during which they did not have an outgroup interaction). However, the models we employed should in principle be resistant to such imbalances.} In support of this possibility, once we used multilevel modeling to separate effects into a between-subjects and within-subjects part, we found that when having an interaction with an outgroup member (vs. not having an outgroup interaction) is a strong predictor of outgroup attitudes across all three studies (even when taking other non-outgroup interactions into account; also see Figure \ref{fig:ContactHypothesis}B). We thus find that the main benefits of intergroup contact remain visible across the three studies. However, the intensive longitudinal data of natural intergroup contacts also offers a novel opportunity to explore how people recall and cognitively aggregate outgroup interactions. Future studies may want to ask participants explicitly to evaluate the number and quality of outgroup interactions following a phase of intensive longitudinal data collection to disentangle the aggregation differences.

% Allport's conditions in ESM data [consistent predictor and partially through quality]
Using the data from our third study, we find that Allport's conditions are related to higher interaction quality perceptions and more positive outgroup attitudes. And when we consider interaction quality and Allport's conditions jointly, we find further evidence that part of the benefits of Allport's conditions are due to the case that such interactions are also perceived as more positive. We thus find first evidence that Allport's conditions of optimal contact are also relevant to the daily interactions recent migrants have in their interactions with majority group members.

% Need fulfillment in intergroup contact [consistent predictor, through quality, slightly better than Allport and SDT]
Finally, when looking at the results for our main proposal regarding the importance of situational key needs, we find highly consistent positive results across all three studies. We find that in all three extensive longitudinal datasets, the fulfillment of core needs during intergroup contacts predicts higher interaction quality perceptions and more positive outgroup attitudes. We find that in all three studies the effect of core need fulfillment on outgroup attitudes goes either fully or partially through interaction quality. We also find that need fulfillment is an important predictor even when taking other fundamental psychological needs or Allport's conditions into account. In fact, our core situational needs measure predicted outgroup attitudes slightly better than Allport's conditions and consistently explained more variance in outgroup attitudes than any of the other psychological needs. In most cases, the core need even took over the variances previously explained by the self-determination theory needs. We thus find strong evidence that within everyday life interactions of recent migrants with majority outgroup members, the perception that one's interaction-specific psychological needs are fulfilled offers a meaningful and flexible predictor of outgroup attitudes.

\subsection{Limitations}
% (1) Sample = minority and (voluntary) migrants, (2) Methodology = short scales (reliability) and no lagged effects, (3) Core Need Concept = non-specific
Among the limitations of our studies we should consider the following. One limitation is our choice of samples. While we believe that a need–based mechanism should be relevant to any inter-group contact, our sample has focused on a minority- and (voluntary) migrant perspective. Without additional evidence it thus remains difficult to judge whether motivational effects will generalize to other migrant groups (e.g., forced migrants), other intergroup contexts (e.g., gender-, religious-, or sexual orientation groups), and to majority groups in their outgroup attitudes. We know of no research that for other contexts psychological needs would not be relevant (once existential needs are met) but future research may extend our findings to other contexts to build an even broader understanding of psychological needs in intergroup contacts. A second limitation lies in our methodology. While extensive longitudinal data is close to real-life events, this method comes at the expense of longer and more robust scales. Long and repetitive scales are often not feasible in extensive longitudinal methods because of the increased burden to the participants. To circumvent this shortcoming, we have ensured that the measures we used were, whenever possible, based on past validations. However, the circumstance remains that extensive longitudinal data often does not allow the same scrutiny of measurement reliability as single–shot cross–section data sets. An additional methodological question lies in the unexplored potential of the longitudinal aspects of our data. For our research questions, we have focused on contemporaneous effects within the data set, yet future investigations should seek to extend the mechanism to developmental trajectories within and between participants. And finally, our conceptualization of core situational needs has been focused on the most essential test of a motivational mechanism. This comes at the expense of specificity in the situational needs (i.e., we have not explored whether different types of needs have stronger effects on interaction quality and outgroup attitudes). With the adaptive measurement we used such an investigation would be possible (e.g., by clustering or classifying the free text entries) but would not have been relevant to our theory–focused research question. Nonetheless, future research may explore which exact needs are most important in different intergroup contexts.

\subsection{Implications}
% ESM data
Given these limitations, we can nonetheless draw a number of implications for other researchers and practitioners --- ranging from the benefits of longitudinal data to theoretical implications. A first implication concerns the feasibility and usefulness of extensive longitudinal data for intergroup contact research and the broader field of social psychology. While setting up an extensive longitudinal study is not easy, we believe the efforts to be similar to a sizable cross-sectional data collection (i.e., for a longitudinal or a high-quality cross-sectional data set with roughly 3,000 intergroup interactions captured and over 10,000 data points in total). Given the comparable effort, this opens up the possibility to explore research questions that focus on real-life phenomena outside the lab or focus on phenomena that depend on changes and influences over time. In the context of intergroup contact research, we are among the first to answer calls to test intergroup contact mechanisms using extended real-life data \citep[e.g.,][]{Pettigrew2011, MacInnis2015}. In doing so, we not only collected an unprecedented amount of real-life data, but our consideration of extensive longitudinal data may present new inconsistencies in how participants perceive and cognitively aggregate their past interactions with other groups — which may suggest large-scale recall-biases or conflations of within and between participant effects in conventional cross-section studies.

% Motivation and core needs 
A broader theoretical implication relates to the role of situational motivation in intergroup contacts. Our results offer a first promising test of psychological needs in intergroup contact. While our results are tentative given their novelty within the field, they were highly consistent across studies and may offer new theoretical avenues. Psychological needs are a facet of the human experience that has thus far been under–emphasized in the intergroup contact literature. This stands in stark contrast of the many cognitive \citep[e.g.,][]{Pettigrew1998, Brown2005} and emotional aspects investigated within the field \citep[e.g.,][]{Stephan2008, Paolini2004}. However, if motivational mechanisms are indeed shown to be meaningful factors in intergroup contacts, future research may be able to integrate broader theoretical frameworks of intergroup contact (e.g., motivations guiding cognition and affect, which in turn drive behavior. cf., theory of reasoned goal pursuit; \citealp{Ajzen2019}. Also see \citealp{Kreienkamp2022d}).

% practical use of needs
Additionally, situational motivations in intergroup contact also offer promising avenues for practitioners and policy-makers. Intergroup contact theory is among the most implemented psychological theories \citep[e.g.,][]{Pettigrew2006, AlRamiah2012a, Reimer2021}. Given our findings that psychological needs in every-day intergroup contacts were at least as powerful as Allport's conditions in predicting outgroup attitudes, considerations of people's needs offer a substantially more immediate mechanism to address. In cases where some or all optimal contact conditions are not possible to be fulfilled, needs offer an even more compelling alternative (e.g., where equal status is contextually not possible or in cases where people help despite a lack of institutional support). Additionally, our conceptualization of situational needs offers a clear opportunity for practitioners and interventions. Instead of addressing needs as a one-size fits all solution (e.g., simply focusing on competence needs), one may at times ask outgroup interaction partners what they need during an interaction. This is not to say that we should not explore which needs tend to be relevant to specific groups in specific intergroup contact contexts. Rather, during interventions for which data on important need contents are not available or infeasible to collect, a flexible and reactive approach of inquiring momentary intergroup contact needs might be more fruitful.

\subsection{Conclusion}
% Conclusion paragraph
In sum, we used extensive longitudinal methodologies to capture real-life interactions of recent migrants with the majority outgroup. Our three studies showcase the feasibility and utility of such data to test intergroup contact theory. We provide evidence that the fulfillment of situational psychological needs during real-life intergroup contacts meaningfully predicts perceived interaction quality and positive outgroup attitudes. Our results point to motivation as an understudied aspect of intergroup contact that is important in understanding when and why an interaction is perceived as positive and will lead to more positive outgroup attitudes.

% Tables
\input{Tables/descrFullWide}
\input{Tables/descrOutWide}

\input{Tables/mdlContactGeneralLong}
\input{Tables/mdlTheorylong}
\input{Tables/mdlRobustlong}

% Figures
\begin{figure}
  \caption{Path Diagrams Situational Needs Model across Studies}
  \label{fig:MainPaths}
  \begin{center}
    \input{PathModels.tex}
  \end{center}
  \caption*{Note: \\
  (A) Study 1; (B) Study 2; (C) Study 3.\\
  General: Path coefficients are standardized regression coefficients. Statistical significance markers are based on the unstandardized regression results (as presented in Table \ref{tab:intergroupNeedsTblLong}).\\
  **** p < .0001, *** p < .001, ** p < .01, * p < .05}
\end{figure}

\begin{figure}
  \caption{Contact Hypothesis}
  \label{fig:ContactHypothesis}
  \centering\includegraphics[width=\textwidth]{Figures/forestParametricREMLGeneralLmer.png}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.75\textwidth]{Figures/forestParametricGeneralLm.png}
  %\end{subfigure}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.75\textwidth]{Figures/forestParametricFEGeneralLmer.png}
  %\end{subfigure}
  \caption*{Note: \\
  Summary of mixed models results of the contemporaneous contact effects.\\
  General: Random effects meta analytic results are presented for completeness.}
\end{figure}

\begin{figure}
  \caption{Core Need Fulfillment}
  \label{fig:AllportNeedFulfillment}
  \centering\includegraphics[width=\textwidth]{Figures/forestParametricTheoryComb.png}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.6\textwidth]{Figures/forestParametricFETheoryQualityCore.png}
  %\end{subfigure}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.6\textwidth]{Figures/forestParametricFETheoryAttitudeCore.png}
  %\end{subfigure}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.6\textwidth]{Figures/forestParametricFETheoryAttitudeCoreQuality.png}
  %\end{subfigure}
  \caption*{Note: \\
  (a) Core Need Fulfillment predicting Interaction Quality.\\
  (b) Core Need Fulfillment predicting Outgroup Attitudes.\\
  (c) Core Need Fulfillment and Interaction Quality predicting Outgroup Attitudes.\\
  General: Random effects meta analytic results are presented for completeness.}
\end{figure}

\begin{figure}
  \caption{Robustness Analyses}
  \label{fig:Robustness}
  \centering\includegraphics[width=\textwidth]{Figures/forestParametricRobustnessComb.png}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.65\textwidth]{Figures/forestParametricFERobustContact.png}
  %\end{subfigure}
  %\begin{subfigure}{\textwidth}
  %  \caption{}
  %  \centering\includegraphics[width=0.65\textwidth]{Figures/forestParametricFERobustSDT.png}
  %\end{subfigure}
  \caption*{Note: \\
  (a) Need Fulfillment and Intergroup Contact predicting Outgroup Attitudes (full sample).\\
  (b) Core Need Fulfillment predicting Outgroup Attitudes while controlling for self-determination theory needs (intergroup contact sample).\\
  General: Random effects meta-analytic results are presented for completeness.}
\end{figure}

\printbibliography

\appendix

\section{Appendix A: Hypotheses}
\label{app:AppendixHypotheses}

\begin{hyp}[H\ref{hyp:contact}] \label{hyp:contact}
Based on the most general understanding of the contact hypothesis, an increase in frequency and quality of contact should jointly account for changes in more favorable outgroup attitudes.
\end{hyp}

\begin{subhyp}[H\ref{hyp:contactFreq}] \label{hyp:contactFreq}
\addtolength{\leftskip}{2.5em}
Participants with more intergroup interactions should have more favorable outgroup attitudes.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:contactDummy}] \label{hyp:contactDummy}
\addtolength{\leftskip}{2.5em}
Outgroup attitudes should be more positive after an intergroup interaction compared to a non-outgroup interaction.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:contactFreqQual}] \label{hyp:contactFreqQual}
\addtolength{\leftskip}{2.5em}
Participants with more intergroup interactions should have more favorable outgroup attitudes depending on the average interaction quality.
\end{subhyp}

\begin{hyp}[H\ref{hyp:AllportsConditions}] \label{hyp:AllportsConditions}
Based on Allport's optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions.
\end{hyp}

\setcounter{subhyp}{0}
\begin{subhyp}[H\ref{hyp:AllportsPred}] \label{hyp:AllportsPred}
\addtolength{\leftskip}{2.5em}
Based on Allport's optimal contact conditions, outgroup attitudes should be more favorable after intergroup interactions with equal status, common goals, collaboration, and structural support.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:AllportsQuality}] \label{hyp:AllportsQuality}
\addtolength{\leftskip}{2.5em}
Based on past research on the role of interaction quality, interaction quality should be more perceived as more favorable after intergroup interactions with equal status, common goals, collaboration, and structural support.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:AllportsQualityMediation}] \label{hyp:AllportsQualityMediation}
\addtolength{\leftskip}{2.5em}
Based on past research on the role of interaction quality, the variance explained in outgroup attitudes by Allport's optimal contact should to a large extent be assumed by interaction quality.
\end{subhyp}

\begin{hyp}[H\ref{hyp:keyNeed}] \label{hyp:keyNeed}
Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions.
\end{hyp}

\setcounter{subhyp}{0}
\begin{subhyp}[H\ref{hyp:keyNeedPred}] \label{hyp:keyNeedPred}
\addtolength{\leftskip}{2.5em}
Outgroup attitudes should be more favorable after intergroup interactions with high key-need fulfillment.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:keyNeedQual}] \label{hyp:keyNeedQual}
\addtolength{\leftskip}{2.5em}
Interaction Quality should be perceived as more positive after intergroup interactions with higher key need fulfillment.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:keyNeedMediation}] \label{hyp:keyNeedMediation}
\addtolength{\leftskip}{2.5em}
The variance explained in outgroup attitudes by key need fulfillment should to a large extend be assumed by interaction quality.
\end{subhyp}

\begin{subhyp}[H\ref{hyp:keyNeedContactType}] \label{hyp:keyNeedContactType}
\addtolength{\leftskip}{2.5em}
The effect of key need fulfillment on outgroup attitudes should be specific to intergroup interactions and not be due to need fulfillment in general. Thus, the effect of key need fulfillment on outgroup attitudes should be stronger for intergroup interactions than for ingroup interactions. 
\end{subhyp}

\begin{subhyp}[H\ref{hyp:keyNeedSDT}] \label{hyp:keyNeedSDT}
\addtolength{\leftskip}{2.5em}
The effect of key need fulfillment on outgroup attitudes should persist even when taking other fundamental psychological needs into account. Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for autonomy, competence, and relatedness fulfillment during the interaction (cf., self-determination theory). 
\end{subhyp}

\begin{hyp}[H\ref{hyp:comparison}] \label{hyp:comparison}
Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict outgroup attitudes at least as well as Allport's conditions.
\end{hyp}

\setcounter{subhyp}{0}
\begin{subhyp}[H\ref{hyp:compModel}] \label{hyp:compModel}
\addtolength{\leftskip}{2.5em}
The need model (H\ref{hyp:keyNeedPred}) should predict more variance in outgroup attitudes than the model based on Allport's conditions (H\ref{hyp:AllportsPred}).
\end{subhyp}

\begin{subhyp}[H\ref{hyp:compTogether}] \label{hyp:compTogether}
\addtolength{\leftskip}{2.5em}
The  effect of key need fulfillment on outgroup attitudes should  persist even when taking other Allport's conditions into account. Thus, the effect of key need fulfillment on outgroup attitudes should remain strong even after controlling for equal status, common goals, collaboration, and structural support.  
\end{subhyp}

\section{Appendix B: Robustness}
\label{app:AppendixRobustness}

\end{document}