### 2022-11-11 PathModels.Rmd
shortened

---
output: latex_fragment
header-includes:
  - \usepackage{tikz}
  - \usepackage{pgfplots}
  - \pgfplotsset{compat=1.17}
  - \tikzset{mynode/.style={draw,text width=1in,align=center} }
  - \usetikzlibrary{positioning}
---

```{r, load_packages, include = FALSE}
lib <- c(
  "rmarkdown",
  "knitr",
  "remedy",
  "bookdown",
  "glue"
)
invisible(lapply(lib, library, character.only = TRUE))
rm(lib)
```

```{r, load_function, include=FALSE}
# \\path (xd.north west) ++(-0.2, 3pt) node[above left]{(A)};
# \\path (xd.north west) ++(-0.4, 16pt) node[draw, above left, shape=circle, very thick, inner sep=2pt] {\\large \\textbf{<<fig_num>>}};
path_tikz_fig <- function(data) {
  glue::glue_data(data, 
"
\\begin{figure}
\\begin{center}
\\begin{tikzpicture}[font=\\sffamily]
    \\node[mynode] (m){<<lab_m>>};
    \\node[mynode,below left=of m](x) {<<lab_x>>};
    \\node[mynode,below right=of m](y) {<<lab_y>>};
    \\node[mynode,above left=of m](xd) {<<lab_x_direct>>};
    \\node[mynode,above right=of m](yd) {<<lab_y_direct>>};
    \\draw[-latex] (xd.east) -- node[above=2mm, align=center] {<<coef_xy_direct>>} (yd.west);
    \\draw[-latex] (x.north) -- node[auto] {<<coef_xm>>} (m.west);
    \\draw[-latex] (m.east) -- node[auto] {<<coef_my>>} (y.north);
    \\draw[-latex] (x.east) -- node[below=2mm, align=center] {<<coef_xy>>} (y.west);
    \\path (xd.north west) ++(-0.2, 6pt) node[above left]{\\textbf{(<<fig_num>>) <<study_num>>}};
\\end{tikzpicture}
\\end{center}
\\end{figure}
", 
.open = "<<", .close = ">>"
)
}

path_tikz <- function(data) {
  glue::glue_data(data, 
"
\\begin{tikzpicture}[font=\\sffamily]
    \\node[mynode] (m){<<lab_m>>};
    \\node[mynode,below left=of m](x) {<<lab_x>>};
    \\node[mynode,below right=of m](y) {<<lab_y>>};
    \\node[mynode,above left=of m](xd) {<<lab_x_direct>>};
    \\node[mynode,above right=of m](yd) {<<lab_y_direct>>};
    \\draw[-latex] (xd.east) -- node[above=2mm, align=center] {<<coef_xy_direct>>} (yd.west);
    \\draw[-latex] (x.north) -- node[auto] {<<coef_xm>>} (m.west);
    \\draw[-latex] (m.east) -- node[auto] {<<coef_my>>} (y.north);
    \\draw[-latex] (x.east) -- node[below=2mm, align=center] {<<coef_xy>>} (y.west);
    \\path (xd.north west) ++(-0.2, 6pt) node[above left]{\\textbf{(<<fig_num>>) <<study_num>>}};
\\end{tikzpicture}
", 
.open = "<<", .close = ">>"
)
}
```

```{r create_diagram, echo = FALSE, results = 'asis'}

load("data/theoryMdlTabl.RData")

y <- "Attitude"
x <- "CoreNeed"
m <- "Quality"
xm <- "CoreNeedQuality"

studies <- c("Study1", "Study2", "Study3")

pathCoeffs <- list()
for (study in studies) {
  mdlDir <- shortMdlPath[[paste(study, y, x, sep = ".")]]
  mdlM <- shortMdlPath[[paste(study, m, x, sep = ".")]]
  mdlFull <- shortMdlPath[[paste(study, y, xm, sep = ".")]]

  pathCoeffs[[study]] <-
    data.frame(
      xy_direct = paste(format(round(mdlDir$estBeta[mdlDir$coef==x],2), nsmall=2), mdlDir$Bstar[mdlDir$coef==x], sep=""), # c
      xm = paste(format(round(mdlM$estBeta[mdlM$coef==x],2), nsmall=2), mdlM$Bstar[mdlM$coef==x], sep=""), # a
      my = paste(format(round(mdlFull$estBeta[mdlFull$coef==m],2), nsmall=2), mdlFull$Bstar[mdlFull$coef==m], sep=""), # b
      xy = paste(format(round(mdlFull$estBeta[mdlFull$coef==x],2), nsmall=2), mdlFull$Bstar[mdlFull$coef==x], sep="") # c'
    )
}

coreNeedPathS1 <-
  data.frame(
    lab_x_direct = "Core Need\\\\Fulfillment",
    lab_y_direct   = "Outgroup\\\\Attitudes",
    lab_x   = "Core Need\\\\Fulfillment",
    lab_m   = "Interaction\\\\Quality",
    lab_y   = "Outgroup\\\\Attitudes",
    coef_xy_direct = pathCoeffs$Study1$xy_direct,
    coef_xm = pathCoeffs$Study1$xm,
    coef_my = pathCoeffs$Study1$my,
    coef_xy = pathCoeffs$Study1$xy,
    fig_num = "A",
    study_num = "Study 1"
  )
coreNeedPathS2 <-
  data.frame(
    lab_x_direct = "Core Need\\\\Fulfillment",
    lab_y_direct   = "Outgroup\\\\Attitudes",
    lab_x   = "Core Need\\\\Fulfillment",
    lab_m   = "Interaction\\\\Quality",
    lab_y   = "Outgroup\\\\Attitudes",
    coef_xy_direct = pathCoeffs$Study2$xy_direct,
    coef_xm = pathCoeffs$Study2$xm,
    coef_my = pathCoeffs$Study2$my,
    coef_xy = pathCoeffs$Study2$xy,
    fig_num = "B",
    study_num = "Study 2"
  )
coreNeedPathS3 <-
  data.frame(
    lab_x_direct = "Core Need\\\\Fulfillment",
    lab_y_direct   = "Outgroup\\\\Attitudes",
    lab_x   = "Core Need\\\\Fulfillment",
    lab_m   = "Interaction\\\\Quality",
    lab_y   = "Outgroup\\\\Attitudes",
    coef_xy_direct = pathCoeffs$Study3$xy_direct,
    coef_xm = pathCoeffs$Study3$xm,
    coef_my = pathCoeffs$Study3$my,
    coef_xy = pathCoeffs$Study3$xy,
    fig_num = "C",
    study_num = "Study 3"
  )

# requires chunk header to be set to results = 'asis'
# cat("\n", path_tikz(coreNeedPathS1), "\n")
# cat("\n", path_tikz(coreNeedPathS2), "\n")
# cat("\n", path_tikz(coreNeedPathS3), "\n")

```

`r path_tikz(coreNeedPathS1)`

\vspace{24pt}

`r path_tikz(coreNeedPathS2)`

\vspace{24pt}

`r path_tikz(coreNeedPathS3)`

<!-- \textit{Note:} Stars based on more conservative unstandardized regressions -->