### 2022-11-11 PathModels.tex

\begin{tikzpicture}[font=\sffamily]
    \node[mynode] (m){Interaction\\Quality};
    \node[mynode,below left=of m](x) {Core Need\\Fulfillment};
    \node[mynode,below right=of m](y) {Outgroup\\Attitudes};
    \node[mynode,above left=of m](xd) {Core Need\\Fulfillment};
    \node[mynode,above right=of m](yd) {Outgroup\\Attitudes};
    \draw[-latex] (xd.east) -- node[above=2mm, align=center] {0.30**} (yd.west);
    \draw[-latex] (x.north) -- node[auto] {0.40****} (m.west);
    \draw[-latex] (m.east) -- node[auto] {0.41***} (y.north);
    \draw[-latex] (x.east) -- node[below=2mm, align=center] {0.13} (y.west);
    \path (xd.north west) ++(-0.2, 6pt) node[above left]{\textbf{(A) Study 1:}};
\end{tikzpicture}

\vspace{24pt}

\begin{tikzpicture}[font=\sffamily]
    \node[mynode] (m){Interaction\\Quality};
    \node[mynode,below left=of m](x) {Core Need\\Fulfillment};
    \node[mynode,below right=of m](y) {Outgroup\\Attitudes};
    \node[mynode,above left=of m](xd) {Core Need\\Fulfillment};
    \node[mynode,above right=of m](yd) {Outgroup\\Attitudes};
    \draw[-latex] (xd.east) -- node[above=2mm, align=center] {0.14***} (yd.west);
    \draw[-latex] (x.north) -- node[auto] {0.27****} (m.west);
    \draw[-latex] (m.east) -- node[auto] {0.24****} (y.north);
    \draw[-latex] (x.east) -- node[below=2mm, align=center] {0.07} (y.west);
    \path (xd.north west) ++(-0.2, 6pt) node[above left]{\textbf{(B) Study 2:}};
\end{tikzpicture}

\vspace{24pt}

\begin{tikzpicture}[font=\sffamily]
    \node[mynode] (m){Interaction\\Quality};
    \node[mynode,below left=of m](x) {Core Need\\Fulfillment};
    \node[mynode,below right=of m](y) {Outgroup\\Attitudes};
    \node[mynode,above left=of m](xd) {Core Need\\Fulfillment};
    \node[mynode,above right=of m](yd) {Outgroup\\Attitudes};
    \draw[-latex] (xd.east) -- node[above=2mm, align=center] {0.18****} (yd.west);
    \draw[-latex] (x.north) -- node[auto] {0.31****} (m.west);
    \draw[-latex] (m.east) -- node[auto] {0.23****} (y.north);
    \draw[-latex] (x.east) -- node[below=2mm, align=center] {0.10**} (y.west);
    \path (xd.north west) ++(-0.2, 6pt) node[above left]{\textbf{(C) Study 3:}};
\end{tikzpicture}