### 2023-05-16 Supplemental Material B - Coding Protocol/margins/FactSheetTheoretical.tex

\input{Supplemental Material A - Coding Protocol/margins/theoNums}

\sffamily
{
\fbox{%
\begin{tabular}{@{}p{\NFwidth}@{}}

% Top Matter
\NFtitle\\
\NFrule
\NFtext{\textbf{Dataset} Theoretical Acculturation Literature}\\
\NFtext{\textbf{Instances Per Dataset} \nTheo\ unique manuscripts}\\

% Motivation Section
\NFRULE
\NFline{Motivation}\\
\NFrule
\NFelement{Original Authors}{}{Kreienkamp, Bringmann, Engler, de Jonge, Epstude}\\
\NFelement{Original Use Case}{}{collection of theoretical works}\\
\NFelement{Original Funding}{}{None}\\

%\NFelement{Additional Field}{}{Info}\\

% Composition Section
\NFRule
\NFline{Composition}\\
\NFrule
\NFelement{Sample or Complete}{}{Sample}\\
\NFelement{Missing Data}{}{\percTheoMissing\% not included}\\
\NFelement{Sensitive Information }{}{author information}\\
%\NFelement{Additional Field}{}{Info}\\

% Collection Process Section
\NFRule
\NFline{Collection}\\
\NFrule
\NFelement{Sampling Strategy}{}{systematic review (see search strategy)}\\
\NFelement{Ethical Review}{}{not applicable}\\
\NFelement{Author Consent}{}{not applicable}\\
%\NFelement{Additional Field}{}{Info}\\

% Cleaning / Labeling Section
\NFRule
\NFline{Cleaning and Labeling}\\
\NFrule
\NFelement{Cleaning Done}{}{Yes}\\
\NFelement{Labeling Done}{}{Yes}\\
%\NFelement{Additional Field}{}{Info}\\

% Uses / Distribution Section
\NFRule
\NFline{Uses and Distribution}\\
\NFrule
\NFelement{Notable Uses}{}{this systematic review}\\
\NFelement{Other Uses}{}{openly accessible for further analysis}\\
\NFelement{Distribution}{}{available on DataVerse and GitHub}\\
%\NFelement{Replicate Distribution}{}{Info}\\
%\NFelement{Additional Field}{}{Info}\\

% Maintenance and Evolution Section
\NFRule
\NFline{Maintenance and Evolution}\\
\NFrule
\NFelement{Corrections or Erratum}{}{None}\\
\NFelement{Methods to Extend}{}{CC BY-NC 4.0}\\
\NFelement{Replicate Maintainers}{}{Jannis Kreienkamp}\\
%\NFelement{Additional Field}{}{Info}\\

\NFRULE
\NFline{Types of theoretical works \hfill\% of manuscripts*}\\
\NFrule
\NFelement{theory}{\nTheories\,manuscripts}{\percTheories\%}\\
\NFrule
\NFelement{framework}{\nFrameworks\,manuscripts}{\percFrameworks\%}\\
\NFrule
\NFelement{model}{\nModels\,manuscripts}{\percModels\%}\\
\NFrule
%\NFelement{Another}{count}{XX\%}\\
%\NFrule
\NFentry{conceptualization \percTheoMissing\%} \\
%\NFentry*{framework 2.9\%}\\
%\NFentry{Red 1.5\%}\\
\NFrule
\NFtext{As identified by the author(s) of the theoretical manuscripts (also see `FrameworkTheoryModel' code).}\\
\NFrule
\NFtext{* Based on final dataset after screening.}
\end{tabular}}

\par} % to end centering