### 2023-05-26 Tables/studentVarDescr.tex

\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:studentVarDescr}Student: Multilevel Core Variable Descriptives}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llccccc}
\toprule
  & Core Need & Competence & Autonomy & Relatedness & Quality & Attitudes NL\\
\midrule
Core Need &  & 0.61*** & 0.66*** & 0.42*** & 0.80*** & 0.10\\
Competence & 0.24*** &  & 0.74*** & 0.66*** & 0.75*** & 0.07\\
Autonomy & 0.15*** & 0.34*** &  & 0.58*** & 0.69*** & -0.04\\
Relatedness & 0.45*** & 0.29*** & 0.42*** &  & 0.55*** & -0.12\\
Quality & 0.17*** & 0.39*** & 0.08*** & 0.05** &  & 0.09\\
Attitudes NL & 0.35*** & 0.43*** & 0.11*** & 0.11*** & 0.13*** & \\
\addlinespace
Grand Mean & 84.87 & 72.55 & 82.59 & 61.21 & 83.77 & 67.26\\
Between SD & 9.17 & 14.47 & 11.21 & 13.36 & 9.12 & 18.64\\
Within SD & 20.33 & 21.17 & 16.06 & 28.74 & 16.80 & 9.40\\
ICC(1) & 0.15 & 0.30 & 0.32 & 0.17 & 0.20 & 0.80\\
ICC(2) & 0.89 & 0.95 & 0.95 & 0.90 & 0.88 & 0.99\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Upper triangle: Between-person correlations;}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Lower triangle: Within-person correlations;}\\
\multicolumn{7}{l}{\rule{0pt}{1em}*** p < .001, ** p < .01,  * p < .05}\\
\end{tabular}}
\end{minipage}
\end{table}