### 2023-05-30 Tables/mdlTheory.tex

\begin{landscape}\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:intergroupNeedsTblWide}Core Need Fulfillment and Allport's Conditions}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccccccccccccccccccccccccc}
\toprule
\multicolumn{1}{c}{} & \multicolumn{6}{c}{Study 1 - CoreNeed} & \multicolumn{6}{c}{Study 2 - CoreNeed} & \multicolumn{6}{c}{Study 3 - CoreNeed} & \multicolumn{6}{c}{Study 3 - Allport} & \multicolumn{2}{c}{Study 3 - CoreNeed + Allport} \\
\cmidrule(l{3pt}r{3pt}){2-7} \cmidrule(l{3pt}r{3pt}){8-13} \cmidrule(l{3pt}r{3pt}){14-19} \cmidrule(l{3pt}r{3pt}){20-25} \cmidrule(l{3pt}r{3pt}){26-27}
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Quality } & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Quality } & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Quality } & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Quality } & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Attitude} & \multicolumn{2}{c}{Attitude} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-15} \cmidrule(l{3pt}r{3pt}){16-17} \cmidrule(l{3pt}r{3pt}){18-19} \cmidrule(l{3pt}r{3pt}){20-21} \cmidrule(l{3pt}r{3pt}){22-23} \cmidrule(l{3pt}r{3pt}){24-25} \cmidrule(l{3pt}r{3pt}){26-27}
 & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$\\
\midrule
\addlinespace[0.3em]
\multicolumn{27}{l}{\textbf{Fixed}}\\
\hspace{1em}(Intercept) & 67.16**** [62.99, 71.34] &  & 72.53**** [66.81, 78.26] &  & 72.50**** [66.74, 78.26] &  & 67.46**** [65.17, 69.76] &  & 70.71**** [67.55, 73.87] &  & 70.69**** [67.52, 73.85] &  & 72.20**** [68.71, 75.70] &  & 68.32**** [65.10, 71.54] &  & 68.32**** [65.10, 71.54] &  & 72.23**** [68.75, 75.72] &  & 68.36**** [65.16, 71.56] &  & 68.36**** [65.14, 71.58] &  & 68.33**** [65.05, 71.61] & \\
\hspace{1em}CoreNeed & 0.37**** [ 0.28,  0.47] & 0.39 [0.30, 0.48] & 0.17** [ 0.06,  0.29] & 0.30 [0.18, 0.41] & 0.05 [-0.05,  0.14] & 0.12 [0.00, 0.24] & 0.29**** [ 0.19,  0.40] & 0.20 [0.10, 0.31] & 0.13*** [ 0.07,  0.19] & 0.14 [0.06, 0.22] & 0.06* [ 0.01,  0.11] & 0.09 [0.01, 0.17] & 0.45**** [ 0.36,  0.54] & 0.31 [0.24, 0.37] & 0.19**** [ 0.12,  0.27] & 0.18 [0.11, 0.24] & 0.12*** [ 0.06,  0.19] & 0.10 [0.04, 0.16] &  &  &  &  &  &  & 0.13**** [ 0.08,  0.17] & 0.12 [0.05, 0.18]\\
\hspace{1em}Quality &  &  &  &  & 0.24*** [ 0.13,  0.35] & 0.45 [0.31, 0.59] &  &  &  &  & 0.16**** [ 0.11,  0.21] & 0.22 [0.13, 0.30] &  &  &  &  & 0.16**** [ 0.11,  0.20] & 0.24 [0.18, 0.31] &  &  &  &  & 0.16**** [ 0.11,  0.21] & 0.23 [0.16, 0.29] &  & \\
\hspace{1em}Allport &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  & 0.62**** [ 0.49,  0.74] & 0.40 [0.33, 0.47] & 0.22**** [ 0.15,  0.29] & 0.21 [0.14, 0.27] & 0.11** [ 0.05,  0.18] & 0.11 [0.05, 0.17] & 0.17*** [ 0.09,  0.24] & 0.17 [0.11, 0.23]\\
\addlinespace[0.3em]
\multicolumn{27}{l}{\textbf{Random}}\\
\hspace{1em}$\sigma^2$ & 224.8 &  & 63.08 &  & 35.4 &  & 267.5 &  & 89.83 &  & 79.34 &  & 230 &  & 91.84 &  & 84.56 &  & 197.5 &  & 93.05 &  & 86.99 &  & 90.78 & \\
\hspace{1em}$\tau_{00}$ & 77.54 (PID) &  & 173.49 (PID) &  & 178.05 (PID) &  & 98.57 (PID) &  & 259.79 (PID) &  & 263.15 (PID) &  & 207.01 (PID) &  & 182.38 (PID) &  & 182.78 (PID) &  & 207.46 (PID) &  & 179.74 (PID) &  & 182.20 (PID) &  & 189.33 (PID) & \\
\hspace{1em}$\tau_{11}$ &  &  & 0.05 (PID.CoreNeed) &  & \makecell[c]{0.03 (PID.CoreNeed)\\0.05 (PID.Quality)} &  & 0.07 (PID.CoreNeed) &  & 0.02 (PID.CoreNeed) &  & \makecell[c]{0.01 (PID.CoreNeed)\\0.02 (PID.Quality)} &  & 0.07 (PID.CoreNeed) &  & 0.05 (PID.CoreNeed) &  & \makecell[c]{0.03 (PID.CoreNeed)\\0.01 (PID.Quality)} &  & 0.17 (PID.Allport) &  & 0.05 (PID.Allport) &  & \makecell[c]{0.02 (PID.Allport)\\0.02 (PID.Quality)} &  & \makecell[c]{0.00 (PID.CoreNeed)\\0.04 (PID.Allport)} & \\
\hspace{1em}$\rho_{01}$ &  &  & 0.03 (cor.slope.intercept) &  & \makecell[c]{ 0.37 (PID.CoreNeed)\\-0.26 (PID.Quality)} &  & 0.27 (cor.slope.intercept) &  & -0.16 (cor.slope.intercept) &  & \makecell[c]{-0.10 (PID.CoreNeed)\\-0.27 (PID.Quality)} &  & 0.06 (cor.slope.intercept) &  & -0.26 (cor.slope.intercept) &  & \makecell[c]{-0.27 (PID.CoreNeed)\\-0.27 (PID.Quality)} &  & 0.24 (cor.slope.intercept) &  & -0.19 (cor.slope.intercept) &  & \makecell[c]{-0.10 (PID.Allport)\\-0.38 (PID.Quality)} &  & \makecell[c]{-1.00 (PID.CoreNeed)\\-0.07 (PID.Allport)} & \\
\hspace{1em}$\rho_{00}$ &  &  &  &  & -0.34 (PID.CoreNeed-Quality) &  &  &  &  &  & 0.74 (PID.CoreNeed-Quality) &  &  &  &  &  & 0.36 (PID.CoreNeed-Quality) &  &  &  &  &  & 0.09 (PID.Allport-Quality) &  & 0.07 (PID.CoreNeed-Allport) & \\
\addlinespace[0.3em]
\multicolumn{27}{l}{\textbf{Fit}}\\
\hspace{1em}name &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  &  & \\
\hspace{1em}AIC & 3240 &  & 2816 &  & 2651 &  & 8062 &  & 7222 &  & 7158 &  & 14129 &  & 12632 &  & 12538 &  & 13918 &  & 12652 &  & 12576 &  & 12625 & \\
\hspace{1em}BIC & 3255 &  & 2839 &  & 2690 &  & 8091 &  & 7251 &  & 7206 &  & 14161 &  & 12665 &  & 12592 &  & 13950 &  & 12684 &  & 12630 &  & 12679 & \\
\hspace{1em}N & 387 &  & 387 &  & 387 &  & 935 &  & 935 &  & 935 &  & 1676 &  & 1672 &  & 1672 &  & 1676 &  & 1672 &  & 1672 &  & 1672 & \\
\hspace{1em}Groups & 21 &  & 21 &  & 21 &  & 108 &  & 108 &  & 108 &  & 70 &  & 70 &  & 70 &  & 70 &  & 70 &  & 70 &  & 70 & \\
\hspace{1em}ICC & 0.2565 &  & 0.7458 &  & 0.8472 &  & 0.3052 &  & 0.7471 &  & 0.7741 &  & 0.485 &  & 0.6738 &  & 0.6943 &  & 0.5399 &  & 0.6679 &  & 0.6863 &  &  & \\
$R^2_{marginal}$ / $R^2_{conditional}$ & 0.103 / 0.333 &  & 0.030 / 0.753 &  & 0.068 / 0.858 &  & 0.054 / 0.343 &  & 0.012 / 0.750 &  & 0.024 / 0.780 &  & 0.063 / 0.517 &  & 0.019 / 0.680 &  & 0.039 / 0.706 &  & 0.111 / 0.591 &  & 0.024 / 0.676 &  & 0.041 / 0.699 &  & 0.086 / NA & \\
\bottomrule
\multicolumn{27}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{27}{l}{\rule{0pt}{1em}PID = Participant ID number (grouping variable)}\\
\multicolumn{27}{l}{\rule{0pt}{1em}$sigma^2$: residual variance, $tau_{00}$: random intercept variance, $tau_{11}$: random slope variance(s), $rho_{01}$: random slope-intercept correlation(s) , $rho_{00}$: correlation(s) between random slopes}\\
\multicolumn{27}{l}{\rule{0pt}{1em}**** $p$ < .0001, *** $p$ < .001, ** $p$ < .01, * $p$ < .05}\\
\end{tabular}}
\end{minipage}
\end{table}
\end{landscape}