### 2023-05-30 Tables/mdlContactGeneral.tex

\begin{landscape}\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:intergroupGeneralTblWide}Intergroup General}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccccccccccc}
\toprule
\multicolumn{1}{c}{} & \multicolumn{6}{c}{Aggregated [across participants]} & \multicolumn{6}{c}{Contemporaneous [within participants]} \\
\cmidrule(l{3pt}r{3pt}){2-7} \cmidrule(l{3pt}r{3pt}){8-13}
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Study 1} & \multicolumn{2}{c}{Study 2} & \multicolumn{2}{c}{Study 3} & \multicolumn{2}{c}{Study 1} & \multicolumn{2}{c}{Study 2} & \multicolumn{2}{c}{Study 3} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13}
 & $B$ & $\eta_p$ & $B$ & $\eta_p$ & $B$ & $\eta_p$ & $B$ & $\beta$ & $B$ & $\beta$ & $B$ & $\beta$\\
\midrule
(Intercept) & 70.79**** [64.32, 77.26] &  & 67.57**** [64.13, 71.01] &  & 64.71**** [61.34, 68.08] &  & 74.12**** [ 56.74, 91.49] &  & 67.05**** [ 57.37, 76.73] &  & 60.36**** [ 46.45, 74.27] & \\
SumContactNL & 0.14 [-0.55,  0.83] & 0.01 & 0.81** [ 0.33,  1.29] & 0.08 & 0.12 [-0.12,  0.36] & 0.02 &  &  &  &  &  & \\
AvQualityComb & 0.35 [-0.44,  1.14] & 0.03 & 0.27 [-0.05,  0.58] & 0.02 & 0.22 [-0.05,  0.49] & 0.04 &  &  &  &  &  & \\
SumContactNL:AvQualityComb & -0.01 [-0.06,  0.03] & 0.02 & 0.02 [-0.03,  0.06] & 0.01 & -0.01 [-0.03,  0.01] & 0.01 &  &  &  &  &  & \\
OutgroupInteraction &  &  &  &  &  &  & 2.48**** [  1.37,  3.59] & 1.07 [ 0.59, 1.54] & 2.83*** [  1.28,  4.38] & 0.11 [ 0.07, 0.16] & 5.57**** [  3.90,  7.23] & 0.21 [ 0.15, 0.27]\\
NonOutgroupInteraction &  &  &  &  &  &  & 0.44 [ -0.67,  1.55] & 0.16 [-0.32, 0.63] & 0.02 [ -0.76,  0.79] & 0.01 [-0.03, 0.05] & 0.36 [ -0.48,  1.19] & 0.02 [-0.02, 0.05]\\
\addlinespace[0.3em]
\multicolumn{13}{l}{\textbf{Random}}\\
\hspace{1em}OutgroupInteractionM &  &  &  &  &  &  & 0.99 [-26.61, 28.59] & 1.65 [-4.27, 7.57] & 26.53** [  8.61, 44.46] & 0.00 [-0.03, 0.03] & 14.14 [ -2.19, 30.48] & 0.00 [-0.03, 0.03]\\
\hspace{1em}NonOutgroupInteractionM &  &  &  &  &  &  & -4.82 [-30.26, 20.63] & -1.08 [-7.02, 4.86] & -8.32 [-22.98,  6.35] & 0.00 [-0.03, 0.03] & -2.75 [-20.00, 14.50] & 0.00 [-0.03, 0.03]\\
\hspace{1em}$\sigma^2$ &  &  &  &  &  &  & 69.91 &  & 77.42 &  & 88.1 & \\
\hspace{1em}$\tau_{00}$ &  &  &  &  &  &  & 179.20 (PID) &  & 318.62 (PID) &  & 199.47 (PID) & \\
\hspace{1em}$\tau_{11}$ &  &  &  &  &  &  &  &  & \makecell[c]{48.96 (PID.OutgroupInteractionC)\\ 7.40 (PID.NonOutgroupInteractionC)} &  & \makecell[c]{40.64 (PID.OutgroupInteractionC)\\ 3.44 (PID.NonOutgroupInteractionC)} & \\
\addlinespace[0.3em]
\multicolumn{13}{l}{\textbf{Fit}}\\
\hspace{1em}$\rho_{01}$ &  &  &  &  &  &  &  &  & \makecell[c]{-0.39 (PID.OutgroupInteractionC)\\ 0.33 (PID.NonOutgroupInteractionC)} &  & \makecell[c]{-0.26 (PID.OutgroupInteractionC)\\ 0.05 (PID.NonOutgroupInteractionC)} & \\
\hspace{1em}$\rho_{00}$ &  &  &  &  &  &  &  &  & 0.03 (PID.OutgroupInteractionC-NonOutgroupInteractionC) &  & 0.24 (PID.OutgroupInteractionC-NonOutgroupInteractionC) & \\
\hspace{1em}Groups &  &  &  &  &  &  & 23 &  & 113 &  & 71 & \\
\hspace{1em}N & 21 &  & 113 &  & 71 &  & 1225 &  & 4965 &  & 3907 & \\
\hspace{1em}F Test & F(3, 17) = 0.40, p =  0.754 &  & F(3, 109) = 4.05, p =  0.009 &  & F(3, 67) = 1.49, p =  0.225 &  &  &  &  &  &  & \\
\hspace{1em}$R^{2 (a)}$ & 0.066 / -0.099 &  & 0.100 / 0.076 &  & 0.063 / 0.021 &  & 0.009 / 0.722 &  & 0.050 / 0.818 &  & 0.049 / 0.717 & \\
\hspace{1em}name &  &  &  &  &  &  &  &  &  &  &  & \\
AIC &  &  &  &  &  &  & 8778 &  & 36461 &  & 29065 & \\
BIC &  &  &  &  &  &  & 8814 &  & 36539 &  & 29140 & \\
ICC &  &  &  &  &  &  & 0.7193 &  & 0.8082 &  & 0.703 & \\
\bottomrule
\multicolumn{13}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{13}{l}{\rule{0pt}{1em}PID = Participant ID number (grouping variable)}\\
\multicolumn{13}{l}{\rule{0pt}{1em}$sigma^2$: residual variance, $tau_{00}$: random intercept variance, $tau_{11}$: random slope variance(s), $rho_{01}$: random slope-intercept correlation(s) , $rho_{00}$: correlation(s) between random slopes}\\
\multicolumn{13}{l}{\rule{0pt}{1em}**** $p$ < .0001, *** $p$ < .001, ** $p$ < .01, * $p$ < .05}\\
\multicolumn{13}{l}{\rule{0pt}{1em}\textsuperscript{a} For linear regression: $R^2 / R^2_{adjusted}$, for multilevel regression: $R^2_{marginal} / R^2_{conditional}$ }\\
\end{tabular}}
\end{minipage}
\end{table}
\end{landscape}