### 2023-05-30 Tables/studentOutVarDescr.tex

\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:studentOutVarDescr}Student: Multilevel Core Variable Descriptives (Outgroup Contact Only)}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llccccc}
\toprule
  & Core Need & Competence & Autonomy & Relatedness & Quality & Attitudes NL\\
\midrule
Core Need &  & 0.44** & 0.49*** & -0.22 & 0.29 & 0.15\\
Competence & 0.23*** &  & 0.58*** & 0.38** & 0.46*** & 0.22\\
Autonomy & 0.12*** & 0.25*** &  & 0.21 & 0.50*** & 0.06\\
Relatedness & 0.49*** & 0.22*** & 0.46*** &  & 0.48*** & -0.10\\
Quality & 0.19*** & 0.38*** & 0.16*** & 0.15*** &  & 0.08\\
Attitudes NL & 0.32*** & 0.36*** & 0.19*** & 0.22*** & 0.29*** & \\
\addlinespace
Grand Mean & 86.86 & 73.23 & 78.58 & 60.30 & 67.08 & 70.41\\
Between SD & 11.20 & 13.95 & 14.07 & 17.35 & 12.54 & 17.13\\
Within SD & 15.87 & 16.81 & 14.24 & 26.14 & 16.54 & 9.87\\
ICC(1) & 0.14 & 0.27 & 0.40 & 0.19 & 0.24 & 0.72\\
ICC(2) & 0.58 & 0.76 & 0.85 & 0.67 & 0.73 & 0.96\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Upper triangle: Between-person correlations;}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Lower triangle: Within-person correlations;}\\
\multicolumn{7}{l}{\rule{0pt}{1em}*** p < .001, ** p < .01,  * p < .05}\\
\end{tabular}}
\end{minipage}
\end{table}