### 2023-05-30 Tables/studentVarDescr.tex

\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:studentVarDescr}Student: Multilevel Core Variable Descriptives}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llccccc}
\toprule
  & Core Need & Competence & Autonomy & Relatedness & Quality & Attitudes NL\\
\midrule
Core Need &  & 0.60*** & 0.66*** & 0.41*** & 0.59*** & 0.11\\
Competence & 0.24*** &  & 0.73*** & 0.66*** & 0.72*** & 0.06\\
Autonomy & 0.16*** & 0.34*** &  & 0.57*** & 0.65*** & -0.05\\
Relatedness & 0.45*** & 0.22*** & 0.52*** &  & 0.66*** & -0.13\\
Quality & 0.17*** & 0.36*** & 0.08*** & 0.05** &  & 0.11\\
Attitudes NL & 0.35*** & 0.42*** & 0.11*** & 0.11*** & 0.09*** & \\
\addlinespace
Grand Mean & 84.87 & 72.55 & 82.59 & 61.21 & 74.51 & 67.26\\
Between SD & 9.17 & 14.47 & 11.21 & 13.36 & 11.24 & 18.64\\
Within SD & 20.33 & 21.17 & 16.06 & 28.74 & 16.59 & 9.40\\
ICC(1) & 0.15 & 0.30 & 0.32 & 0.17 & 0.29 & 0.80\\
ICC(2) & 0.89 & 0.95 & 0.95 & 0.90 & 0.92 & 0.99\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Upper triangle: Between-person correlations;}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Lower triangle: Within-person correlations;}\\
\multicolumn{7}{l}{\rule{0pt}{1em}*** p < .001, ** p < .01,  * p < .05}\\
\end{tabular}}
\end{minipage}
\end{table}