### 2023-05-30 Tables/workerVarDescr.tex

\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:workerVarDescr}Worker: Multilevel Core Variable Descriptives}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llccccc}
\toprule
  & Core Need & Competence & Autonomy & Relatedness & Quality & Attitudes NL\\
\midrule
Core Need &  & 0.62*** & 0.49** & 0.00 & 0.33 & -0.20\\
Competence & 0.37*** &  & 0.69*** & -0.14 & 0.25 & -0.21\\
Autonomy & 0.32*** & 0.09** &  & -0.04 & 0.21 & 0.19\\
Relatedness & 0.45*** & 0.37*** & 0.35*** &  & 0.09 & -0.06\\
Quality & 0.05 & 0.42*** & 0.11*** & 0.12*** &  & 0.14\\
Attitudes NL & 0.10*** & 0.31*** & 0.15*** & 0.14*** & 0.44*** & \\
\addlinespace
Grand Mean & 77.95 & 62.10 & 72.17 & 55.29 & 67.00 & 71.49\\
Between SD & 14.68 & 13.72 & 12.09 & 14.59 & 9.26 & 12.91\\
Within SD & 20.83 & 20.89 & 15.15 & 23.29 & 17.43 & 8.11\\
ICC(1) & 0.29 & 0.28 & 0.38 & 0.28 & 0.23 & 0.70\\
ICC(2) & 0.96 & 0.95 & 0.97 & 0.95 & 0.83 & 0.99\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Upper triangle: Between-person correlations;}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Lower triangle: Within-person correlations;}\\
\multicolumn{7}{l}{\rule{0pt}{1em}*** p < .001, ** p < .01,  * p < .05}\\
\end{tabular}}
\end{minipage}
\end{table}