### 2023-05-30 Tables/studentOutVarDescr.tex
descr tbl red

\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:studentOutVarDescr}Student: Multilevel Core Variable Descriptives (Outgroup Contact Only)}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llcc}
\toprule
  & Core Need & Quality & Attitudes NL\\
\midrule
Core Need &  & 0.23 & 0.17\\
Quality & 0.23*** &  & 0.06\\
Attitudes NL & 0.16*** & 0.29*** & \\
\addlinespace
Grand Mean & 86.86 & 67.08 & 70.41\\
Between SD & 11.20 & 12.54 & 17.13\\
Within SD & 15.87 & 16.54 & 9.87\\
\addlinespace
ICC(1) & 0.14 & 0.24 & 0.72\\
ICC(2) & 0.58 & 0.73 & 0.96\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{4}{l}{\rule{0pt}{1em}Upper triangle: Between-person correlations;}\\
\multicolumn{4}{l}{\rule{0pt}{1em}Lower triangle: Within-person correlations;}\\
\multicolumn{4}{l}{\rule{0pt}{1em}*** p < .001, ** p < .01,  * p < .05}\\
\end{tabular}}
\end{minipage}
\end{table}