### 2023-05-30 Tables/studentVarDescr.tex

\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:studentVarDescr}Student: Multilevel Core Variable Descriptives}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llcc}
\toprule
  & Core Need & Quality & Attitudes NL\\
\midrule
Core Need &  & 0.61*** & 0.03\\
Quality & 0.21*** &  & 0.02\\
Attitudes NL & 0.08*** & 0.09*** & \\
\addlinespace
Grand Mean & 84.87 & 74.51 & 67.26\\
Between SD & 9.17 & 11.24 & 18.64\\
Within SD & 20.33 & 16.59 & 9.40\\
\addlinespace
ICC(1) & 0.15 & 0.29 & 0.80\\
ICC(2) & 0.89 & 0.92 & 0.99\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{4}{l}{\rule{0pt}{1em}Upper triangle: Between-person correlations;}\\
\multicolumn{4}{l}{\rule{0pt}{1em}Lower triangle: Within-person correlations;}\\
\multicolumn{4}{l}{\rule{0pt}{1em}*** p < .001, ** p < .01,  * p < .05}\\
\end{tabular}}
\end{minipage}
\end{table}