### 2023-05-30 Tables/workerVarDescr.tex

\begin{table}
\begin{minipage}[t][\textheight][t]{\textwidth}

\caption{\label{tab:workerVarDescr}Worker: Multilevel Core Variable Descriptives}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llcc}
\toprule
  & Core Need & Quality & Attitudes NL\\
\midrule
Core Need &  & 0.46* & -0.18\\
Quality & 0.37*** &  & 0.19\\
Attitudes NL & 0.11*** & 0.50*** & \\
\addlinespace
Grand Mean & 77.95 & 67.00 & 71.49\\
Between SD & 14.68 & 9.26 & 12.91\\
Within SD & 20.83 & 17.43 & 8.11\\
\addlinespace
ICC(1) & 0.29 & 0.23 & 0.70\\
ICC(2) & 0.96 & 0.83 & 0.99\\
\bottomrule
\multicolumn{4}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{4}{l}{\rule{0pt}{1em}Upper triangle: Between-person correlations;}\\
\multicolumn{4}{l}{\rule{0pt}{1em}Lower triangle: Within-person correlations;}\\
\multicolumn{4}{l}{\rule{0pt}{1em}*** p < .001, ** p < .01,  * p < .05}\\
\end{tabular}}
\end{minipage}
\end{table}