### 2023-06-01 Supplemental Material D - Coding Protocol/margins/FactSheet.tex
SI D masked

\sffamily
{
\fbox{%
\begin{tabular}{@{}p{\NFwidth}@{}}

% Top Matter
\NFtitle\\
\NFrule
\NFtext{\textbf{Dataset} Acculturation over Time Studies}\\
\NFtext{\textbf{Instances Per Dataset} 2,995 free text responses in 47 topics}\\

% Motivation Section
\NFRULE
\NFline{Motivation}\\
\NFrule
\NFelement{Original Authors}{}{Kreienkamp, Agostini, Bringmann, de Jonge, Epstude}\\
\NFelement{Original Use Case}{}{cultural adaptation in daily interactions}\\
\NFelement{Original Funding}{}{No specific}\\

%\NFelement{Additional Field}{}{Info}\\

% Composition Section
\NFRule
\NFline{Composition}\\
\NFrule
\NFelement{Sample or Complete}{}{Complete}\\
\NFelement{Missing Data}{}{10.33\% excluded}\\
\NFelement{Sensitive Information }{}{free text responses}\\
%\NFelement{Additional Field}{}{Info}\\

% Collection Process Section
\NFRule
\NFline{Collection}\\
\NFrule
\NFelement{Sampling Strategy}{}{experience sampling method (ESM)}\\
\NFelement{Ethical Review}{}{University masked for Peer review}\\
\NFelement{Author Consent}{}{not applicable}\\
%\NFelement{Additional Field}{}{Info}\\

% Cleaning / Labeling Section
\NFRule
\NFline{Cleaning and Labeling}\\
\NFrule
\NFelement{Cleaning Done}{}{Yes}\\
\NFelement{Labeling Done}{}{Yes}\\
%\NFelement{Additional Field}{}{Info}\\

% Uses / Distribution Section
\NFRule
\NFline{Uses and Distribution}\\
\NFrule
\NFelement{Notable Uses}{}{non published data}\\
\NFelement{Other Uses}{}{partially open for further analysis}\\
\NFelement{Distribution}{}{available on OSF and GitHub}\\
%\NFelement{Replicate Distribution}{}{Info}\\
%\NFelement{Additional Field}{}{Info}\\

% Maintenance and Evolution Section
\NFRule
\NFline{Maintenance and Evolution}\\
\NFrule
\NFelement{Corrections or Erratum}{}{None}\\
\NFelement{Methods to Extend}{}{CC BY-NC 4.0}\\
\NFelement{Replicate Maintainers}{}{Jannis Kreienkamp}\\
%\NFelement{Additional Field}{}{Info}\\

\NFRULE
\NFline{Sample Sizes per study \hfill\% of total*}\\
\NFrule
\NFelement{Study 1}{387 interactions of 1,225 measurements}{31.59\%}\\
\NFrule
\NFelement{Study 2}{931 interactions of 4,965 measurements}{18.75\%}\\
\NFrule
\NFelement{Study 3}{1,677 interactions of 4,107 measurements}{40.83\%}\\
% \NFrule
%\NFelement{Another}{count}{XX\%}\\
%\NFrule
% \NFentry{we focus on outgroup interactions for this analysis} \\
%\NFentry*{framework 2.9\%}\\
%\NFentry{Red 1.5\%}\\
\NFrule
\NFtext{For the analysis using this coding we focused on outgroup interactions only (also see `introduction').}\\
\NFrule
\NFtext{* based on outgroup interaction responses.}
\end{tabular}}

\par} % to end centering